---
title: "Development actions history"
output: html_document
editor_options: 
  chunk_output_type: console
---

All commands that you use to use when developing packages...

# First time just after creating the project

- Fill the following chunk to create the DESCRIPTION of your package

```{r description}
# Describe your package
fusen::fill_description(
  overwrite = T,
  pkg = here::here(),
  fields = list(
    Title = "Run ISR analyses for Taxon Profiles",
    Description = "Survival, Reproduction and Growth analyses",
    `Authors@R` = c(
      person("Floriane", "Plard", email = "floriane.c.plard@gmail.com", role = c("aut", "cre", "ctr")),
      person("Species", "360", email = "Support@Species360.org", role = c("cph")),
      person("Fernando", "Colchero",  role = c("ctb"))
    )
  )
)
# Define License with use_*_license()
usethis::use_proprietary_license("Species360")

# ADD THAT TO DESCRUIPTION TO USE paramDemo...
#Remotes: git::https://github.com/fercol/paramDemo/pkg.git
```

# All-in-one function to deploy publicly on GitHub

Either use this GitHub all-in-one function or choose the steps one-by-one in the following sections.
See `vignette("share-on-a-github-website", package = "fusen")` for more information.

```{r, eval=FALSE}
# _All-in-one share of your package and its documentation on GitHub
fusen::init_share_on_github()
```


# Start using git

```{r, eval=FALSE}
usethis::use_git()
# Deal with classical files to ignore
usethis::git_vaccinate()
# Use main for primary branch
usethis::git_default_branch_rename()
```

# Set extra sources of documentation

```{r, eval=FALSE}
# Install a first time
remotes::install_local()
# README
usethis::use_readme_rmd()
# Code of Conduct
usethis::use_code_of_conduct("Support@Species360.org")
# NEWS
usethis::use_news_md()
```

**From now, you will need to "inflate" your package at least once to be able to use the following commands. Let's go to your flat template, and come back here later if/when needed.**


# Package development tools
## Use once

```{r, eval=FALSE}
# Pipe
usethis::use_pipe()

# package-level documentation
usethis::use_package_doc()
```

## Use everytime needed

```{r, eval=FALSE}
# Add a new flat template
fusen::add_flat_template("add")

#inflate without check
fusen::inflate_all_no_check()

# Simulate package installation
pkgload::load_all()

# Generate documentation and deal with dependencies
attachment::att_amend_desc()

# Check the package
devtools::check()

pkgdown::build_site(override = list(destination = "inst/site"))
pkgdown::build_site()
 # build the tar.gz with vignettes to share with others
devtools::build(vignettes = TRUE)
```

## Code coverage
```{r, eval=FALSE}
library(covr)
cov = package_coverage()
zero_coverage(cov)


cov2 =package_coverage(line_exclusions = list("R/make_summary.R", "R/tx_devout.R"))
cov2
zero_coverage(cov2)
```

# Share the package

```{r, eval=FALSE}
# set and try pkgdown documentation website locally
usethis::use_pkgdown()
pkgdown::build_site()
#for some parts
pkgdown::build_home()
pkgdown::build_articles()
pkgdown:::build_reference()
pkgdown::build_news()

# Share your package and its documentation on GitHub
# usethis::create_github_token()
# gitcreds::gitcreds_set()
fusen::init_share_on_github()
```

