---
title: "Reproduction"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Rep_main

```{r function-Rep_main}
#' my_fun Title
#'
#' @return 1
#' @export
#'
#' @examples
# Rep_main <- function() {
#   # Find institutions and their hemisphere:
#   Inst <- newcore$firstInst
#   idinst <- which(as.character(.institution$AnonInstitutionID) %in% Inst)
#   subinst <- .institution[idinst, ]
#   rownames(subinst) <- as.character(subinst$AnonInstitutionID)
#   subinst <- subinst[Inst, ]  
#   
#   # Find which are from the northern and southern hemisphere:
#   idnorth <- which(subinst$Latitude > 0)
#   idsouth <- which(subinst$Latitude < 0)
#   Hemisph <- rep(NA, nrow(subinst))
#   Hemisph[idnorth] <- "North"
#   Hemisph[idsouth] <- "South"
#   instHemisph <- data.frame(AnonInstID = subinst$AnonInstitutionID, 
#                             Continent = subinst$Continent, 
#                             Country = subinst$Country, 
#                             Hemisph = Hemisph)
#   
#   
#   # Extract contraception method:
#   idIncont <- which(as.character(.contraception$` AnonID`) %in% 
#                       as.character(newcore$anonID))
#   subContra <- .contraception[idIncont, ]
#   nContra <- length(idIncont)
#   method <- rep("NA", nContra)
#   genMethod <- c("hormonal", "surgical", "immunological", "Management", 
#                  "Indeterminate", "Undetermined")
#   methodTransl <- c("Hormonal", "Surgical", "Immunological", "Management", 
#                     "Undetermined", "Undetermined")
#   nMeth <- length(genMethod)
#   for (ii in 1:nMeth) {
#     idm <- grep(genMethod[ii], subContra$Method)
#     method[idm] <- methodTransl[ii]
#   }
#   genStatus <- c("Inactive", "Active", "Undetermined", "Indeterminate")
#   statusTransl <- c("Inactive", "Active", "Undetermined", "Undetermined")
#   status <- rep(NA, nContra)
#   for (ii in 1:nMeth) {
#     idm <- grep(genStatus[ii], subContra$Status)
#     status[idm] <- statusTransl[ii]
#   }
#   
#   
#   # Final contraception table:
#   newContra <- data.frame(anonID = as.character(.contraception$` AnonID`)[idIncont],
#                           date = as.Date(.contraception$Date)[idIncont], 
#                           method = method, status = status)
#   
#   if(length(unique(subpar$currentInst))>1){
#     if (fertSumm$NOffsp >= minNrepro) {
#       if (fertSumm$NParent >= minNparepro) {
#         
#         #remove curretn inst to avoid duplicated lines
#         subparfert <- subpar%>%
#           select(-currentInst)%>%
#           distinct()
#         
#         
#         REp_agemat
#         
#         
#       }}}
#   
#   
#   
# }
```

```{r examples-Rep_main}

```

```{r tests-Rep_main}
test_that("Rep_main works", {
  
})
```




# Age at first reproduction

`Rep_agemat()` returns summary statistics for ages of reproduction. It uses the following argument:
* Reprodata, a data frame including at least the columns *ParentAnonID* and *ageBirth*

```{r function-Rep_agemat}
#' Age of reproduction
#' 
#' Returns summary statistcs for age of reproduction
#'
#' @param Reprodata \code{data frame} including at least the columns *ParentAnonID* and *ageBirth*
#'
#' @return A data frame including:
#' * N_moth_agemat: Number of mothers used
#' * N_birth_agemat : Number of births used
#' * ageMat: age at first reproduction (i.e. when 2.5% of females have reproduced for the first time)
#' * ageYouRep: Minimum age at first reproduction
#' * ageOld1Rep: Maximum age at first reproduction
#' * ageMean1Rep: Mean age at first reproduction
#' * ageSd1Rep: Standard deviation of age at first reproduction
#' * ageOldRep: Maximum age of reproduction
#' * ageMeanRep: Mean age of reproduction 
#' * ageSdRep: Standard deviation of age of reproduction 
#' * ageMedRep: Median age of reproduction 
#' * ageMeanRepIC: 95% confident interval of mean age of reproduction 
#' * ageMean1RepIC: 95% confident interval of mean age at first reproduction 
#' 
#' @export
#' @importFrom stats dist median var sd
#'
#' @examples
Rep_agemat <- function(Reprodata) {
  
   assert_that(is.data.frame(Reprodata))
  
  assert_that(Reprodata %has_name% c("ParentAnonID", "ageBirth"))

  
  fertSumm <- tibble(N_moth_agemat = 0,N_birth_agemat = 0, 
                     ageMat = NA,   
                     ageYouRep = NA, ageOldRep = NA, 
                     ageMean1Rep = NA,ageSd1Rep = NA, ageOld1Rep = NA,
                     ageMeanRep = NA, ageSdRep= NA, ageMedRep = NA,
                     ageMeanRepIC = NA, ageMean1RepIC = NA) 
  
  # Age at first reproduction:
  TageMat <-  Reprodata%>%select(ParentAnonID, ageBirth)%>%
    group_by(ParentAnonID)%>%
    summarize(ageMat = min(ageBirth),
              ageold = max(ageBirth))%>%
    ungroup()
    
    fertSumm$N_moth_agemat  = nrow(TageMat)
  fertSumm$N_birth_agemat  = nrow( Reprodata)
  fertSumm$ageMat<- as.numeric(quantile(TageMat$ageMat, 0.025, 
                                              na.rm = TRUE))
  
  fertSumm$ageYouRep= as.numeric(min(TageMat$ageMat, na.rm = TRUE))
  fertSumm$ageOldRep= as.numeric(max( TageMat$ageold, na.rm = TRUE))
  fertSumm$ageOld1Rep = as.numeric(max(TageMat$ageMat, na.rm = TRUE))
  
  fertSumm$ageMean1Rep = mean(as.numeric(TageMat$ageMat), na.rm = TRUE)
  fertSumm$ageSd1Rep = sd(as.numeric(TageMat$ageMat), na.rm = TRUE)
  fertSumm$ageMean1RepIC = list(fertSumm$ageMean1Rep +c(-1.96, 1.96)*sqrt(var(as.numeric(TageMat$ageMat), na.rm = TRUE)/nrow(TageMat)))
  fertSumm$ageMeanRep = mean(as.numeric(Reprodata$ageBirth), na.rm = TRUE)
  fertSumm$ageSdRep = sd(as.numeric(Reprodata$ageBirth), na.rm = TRUE)
  fertSumm$ageMeanRepIC = list(fertSumm$ageMeanRep +c(-1.96, 1.96)*sqrt(var(as.numeric( Reprodata$ageBirth), na.rm = TRUE)/nrow(Reprodata)))
  
  fertSumm$ageMedRep =  median(as.numeric(Reprodata$ageBirth), na.rm = TRUE)
  
  return(fertSumm)
}
```

```{r examples-Rep_agemat}
data(core)
data(collection)
data(parent)
data(moves)
#prepare Data
Data <- Rep_prepdata (coresubset = core, collection, parent, moves)
#Calculate summary for reproductive ages
out <- Rep_agemat(Data$Reprodata)
out

```

```{r tests-Rep_agemat}
test_that("Rep_agemat works", {
  data(core)
data(collection)
data(parent)
data(moves)
Data <- Rep_prepdata (coresubset = core, collection, parent, moves)
out <- Rep_agemat(Data$Reprodata)

expect_named(out, c("N_moth_agemat", "N_birth_agemat", "ageMat", "ageYouRep",
                    "ageOldRep", "ageMean1Rep", "ageSd1Rep", "ageOld1Rep",
                    "ageMeanRep", "ageSdRep", "ageMedRep", "ageMeanRepIC",
                    "ageMean1RepIC"))
expect_true(is.numeric(as.numeric(out[1:11])))
expect_true(is.list(out[12]))
expect_true(is.list(out[13]))
})
```




# litter size

```{r function-Rep_littersize}
#' my_fun Title
#'
#' @return 1
#' @export
#'
#' @examples
Rep_littersize <- function() {
  littSumm <- list(NOffsp_prob = NULL,NParent_prob = NULL, 
                   NReprEvent = NULL,
                   analyzed = FALSE,error = "", Nerr= 0,
                   MeanLittSize = NULL, MedLittSize = NULL)
  
  subpar <- subpar %>%
    filter(Probability > parentProb)
  littSumm$NOffsp_prob <- Nrepros <- length(unique(subpar$AnonID))
  unParentID <- unique(subpar$ParentAnonID)
  littSumm$NParent_prob <- Nparent <- length(unParentID)
  
  if (sx == "Female" | (sx == "All" & !forceBySex[taxa])) {
    
    if(length(unique(subpar$currentInst))>1){
      
      #remove curretn inst to avoid duplicated lines
      subpar <- subpar%>%
        select(-currentInst)%>%
        distinct()
      if (nrow(subpar)>= minNrepro)  {
        
        nindVec <- c()
        parVec <- c()
        parAgeVec <- c()
        for (ip in 1:Nparent) {
          idp <- which(subpar$ParentAnonID == unParentID[ip])
          nIdp <- length(idp)
          ibAge <-as.numeric(subpar$Offspring_BirthDate[idp] - subpar$Parent_BirthDate[idp])
          ibSort <- sort(ibAge)
          ibDiff <- as.numeric(diff(ibSort))
          idNew <- which(ibDiff > 7)
          repEv <- rep(0, nIdp)
          repEv[1] <- 1
          repEv[-1][idNew] <- 1
          nRepEvs <- as.numeric(table(cumsum(repEv)))
          parAgeVec <- c(parAgeVec, ibSort[which(repEv == 1)] / 365.25)
          nindVec <- c(nindVec, c(nRepEvs))
          parVec <- c(parVec, rep(unParentID[ip], length(nRepEvs)))
        }
        littSizeDf <- data.frame(parentID = parVec, age = parAgeVec, 
                                 Noffsp = nindVec)
        unNind <- min(nindVec):max(nindVec)
        nind <- length(unNind)
        littSize <- matrix(NA, nind, 3, 
                           dimnames = list(NULL, c("littSize", "total", 
                                                   "prop")))
        littSize[, "littSize"] <- unNind
        for (iof in 1:nind) {
          littSize[iof, "total"] <- length(which(nindVec == unNind[iof]))
        }
        littSize[, "prop"] <- littSize[, "total"] / sum(littSize[, "total"], 
                                                        na.rm = TRUE)
        # Summary
        littSumm$MeanLittSize =  mean(nindVec)
        littSumm$MedLittSize = median(nindVec)
        littSumm$analyzed = TRUE 
        littSumm$NReprEvent= sum(littSize[, "total"])
      }
    }
  }
}
```

```{r examples-Rep_littersize}

```

```{r tests-Rep_littersize}
test_that("Rep_littersize works", {
  
})
```




# fertility

```{r function-Rep_fert}
#' my_fun Title
#'
#' @return 1
#' @export
#'
#' @examples
Rep_fert <- function() {
  fertSumm <- list(NOffsp_age = 0,  NParent_age = 0,
                   NOffsp_bd = 0,NParent_bd = 0,
                   NOffsp_glob = 0,NParent_glob = 0,
                   NOffsp = 0, NParent = 0, 
                   NAdult = 0, NAdult_rep = 0, 
                   analyzed = FALSE, 
                   Nerr= 0, error = "",
                   Prepmean = NULL , Fertmean = NULL) 
  
  
  
  
  
  
  # Store summary:
  fertSumm$analyzed = TRUE
  
  # Reproduction matrix:
  ageFert <- matrix(0, length(ageVec), 15, 
                    dimnames = list(NULL, c("Age", "NReprAdult", "NtotParent", 
                                            "Prep", "SEprep","Plower","Pupper",
                                            "Fert", "SEfert","Fertlower","Fertupper",
                                            "SexRatio", "SEsexratio","SRlower","SRupper")))
  ageFert[, "Age"] <- ageVec
  Nadstot= Npartot = 0
  for (ix in 1:length(ageVec)) {
    xx <- ageVec[ix]
    Nads <- nrow(ADULTS%>%
                   filter(Age == xx,
                          contra == FALSE))
    
    subage<- subparfert%>% filter(ageBirthDisc == xx)
    Npar = length(unique(subage$ParentAnonID))
    Npartot = Npartot+Npar
    if (Nads > 0 & Npar >0) {
      tot <- subage %>% group_by(ParentAnonID)%>% 
        summarize (rep = max(Probability)/100)
      prepr <- sum(tot$rep) / Nads
      Nadstot= Nadstot+Nads
      fertRate <- sum(subage$Probability)/100 / Npar
      temp = subage%>% group_by(ParentAnonID)%>%summarize(A = sum(Probability)/100, N = n()^2)
      Vfert = sum(temp$A*temp$N) / Npar - fertRate^2
      sr <- length(which(subage$Sex == "Female"))/ length(which(subage$Sex  %in% c("Female", "Male")))
      if (fertRate > 1) {
        if(Vfert>0){
          fertSE <-sqrt(Vfert / Npar)
        }else{ fertSE <- 0}
      } else {
        fertSE <- sqrt(fertRate * (1 - fertRate) / 
                         Npar)
      }
      pSE <-  sqrt(prepr * (1 - prepr) / 
                     Nads)
      if(prepr> 1 | prepr<0){stop(glue("p(repro) = {pSE} for age {xx} and sex {sx}.As a probability, it should be include between 0 and 1"))}
      srSE <-  sqrt( sr * (1 -  sr) / 
                       length(which(subage$Sex  %in% c("Female", "Male"))))
      
      
    } else {
      fertRate <- prepr <- sr <- 0
      pSE <- fertSE <- srSE <- 0
    }
    fertlower = fertRate - 1.96*fertSE
    fertupper = fertRate + 1.96*fertSE
    plower = prepr -1.96 *  pSE
    pupper = prepr +1.96 *  pSE
    srlower = sr -1.96 *  srSE
    srupper = sr +1.96 *  srSE
    
    srlower = ifelse(srlower<0,0,srlower)
    plower = ifelse(plower<0,0,plower)
    fertlower = ifelse(fertlower<0,0,fertlower)
    srupper = ifelse(srupper>1,1,srupper)
    pupper = ifelse(pupper>1,1,pupper)
    ageFert[ix, -1] <- c(Nads, Npar, 
                         prepr, pSE,plower,pupper,
                         fertRate, fertSE,fertlower,fertupper, 
                         sr, srSE,srlower,srupper)
  }
  tot <- subparfert %>% group_by(ParentAnonID,ageBirthDisc)%>% 
    summarize ( rep = max(Probability)/100)
  fertSumm$Prepmean =sum(tot$rep) / Nadstot
  fertSumm$Fertmean = sum(subpar$Probability)/ Npartot /100
  
  
  
}
```

```{r examples-Rep_fert}

```

```{r tests-Rep_fert}
test_that("Rep_fert works", {
  
})
```



# Prepare data for reproduction

`Rep_prepdata()` prepares reproduction data including age of parent at birth, offspring sex and birth date and data of potential reproductive individuals at each age. It uses at least the following arguments:

* cleaned `coresubset` including only the selected individuals. It must includes at least the following columns: *anonID*, 'birthType', *MaxBirthDate*, *MinBirthDate*, *DepartDate*, *BirthDate*,*Sex*, and *firstInst*
* `collection`data including at least the following columns: *AnimalID*, *ScopeType*, *ChangeDate*
* `parent` data including at least the following columns: *ParentAnonID*, *ParentCollectionScopeType*, *OffspringCollectionScopeType*, *AnonID*, *ParentOriginType*, *Probability*
* `moves` data including at least the following columns: *AnonID*, *To*, *Date*


```{r function-Rep_prepdata}
#' Prepare Reproduction date
#' 
#' Prepare reproduction data including age of parent at birth, offspring sex and birth date and data of potential reproductive individuals at each age.
#'
#' @param coresubset \code{data frame} cleaned core data including only the selected individuals. It must includes at least the following columns: *anonID*, 'birthType', *MaxBirthDate*, *MinBirthDate*, *DepartDate*, *BirthDate*,*Sex*, and *firstInst*
#' @param collection \code{data frame} Collection data including at least the following columns: *AnimalID*, *ScopeType*, *ChangeDate*
#' @param parent \code{data frame} Parent data including at least the following columns: *ParentAnonID*, *ParentCollectionScopeType*, *OffspringCollectionScopeType*, *AnonID*, *ParentOriginType*, *Probability*
#' @param moves \code{data frame} Moves data including at least the following columns: *AnonID*, *To*, *Date*
#' @param BirthType_parent \code{character} Captive, Wild, or All. Default =  "Captive"
#' @param BirthType_offspring \code{character} Captive, Wild, or All. Default =  "Captive"
#' @param Age_uncert \code{numeric} Maximum uncertainty (= MaxBirthDate - MinBirthDate) in birth date in days. Default = 10000
#' @param Global \code{logical} Whether only individuals belonging to global collections should be used. Default = TRUE
#'
#' @return A list including
#' * The reproduction data
#' * The potential reproductive Adults
#' * A summary including:
#'  - Nbirths: the raw number of births
#'  - Nadults: the raw number of potential reproductive adults
#'  - NOffsp: the number of offspring with known birth date and with parents with known age
#'  - NParent: the number of unique parents with known age and with offspring with known birth date
#'  - NOffsp_age: the number of offspring with known birth date and with parents with a known age > 0
#'  - NParent_age: the number of unique parents with known and positive age and with offspring with known birth date
#' 
#' @export
#'
#' @examples
Rep_prepdata <- function(coresubset, collection, parent, moves,
                         BirthType_parent = "Captive", BirthType_offspring = "Captive", 
                         Age_uncert = 1000000, Global = TRUE
) {
  
  assert_that(is.data.frame(coresubset))
  assert_that(is.data.frame(collection))
  assert_that(is.data.frame(parent))
  assert_that(is.data.frame(moves))
  
  assert_that(coresubset %has_name% c("anonID", 'birthType', "MaxBirthDate", 
                                      "MinBirthDate", "DepartDate", "BirthDate",
                                      "Sex", "firstInst"))
  assert_that(collection %has_name% c("AnimalID", "ScopeType", "ChangeDate"))
  assert_that(parent %has_name% c("ParentAnonID", "ParentCollectionScopeType", 
                                  "OffspringCollectionScopeType", "AnonID",
                                  "ParentOriginType", "Probability"))
  assert_that(moves %has_name% c("AnonID", "To", "Date"))
  
  assert_that(is.character( BirthType_parent))
  assert_that(BirthType_parent %in% c("Captive", "Wild", "All"))
  assert_that(is.character( BirthType_offspring))
  assert_that( BirthType_offspring %in% c("Captive", "Wild", "All"))
  assert_that(is.numeric(Age_uncert))
  assert_that(is.logical(Global))
  
  fertSumm <- tibble(Nbirths = 0,
                     Nadults = 0,
                     NOffsp = 0, NParent = 0,
                     NOffsp_age = 0, NParent_age = 0
  )
  
  #Offspring               
  if (BirthType_offspring != "All"){
    offspSub <- coresubset %>%
      filter(stringr::str_detect(birthType, pattern = BirthType_offspring))
  }else{offspSub<- coresubset}
  # Number of birth records:
  fertSumm$Nbirths <- nrow(offspSub)
  
  
  # Adults:
  if (BirthType_parent != "All"){
    ADULTS <- core %>%
      filter(stringr::str_detect(birthType, pattern = BirthType_parent))
  }else{ADULTS<- core}
  
  ADULTS <- ADULTS%>%
    filter(as.numeric(MaxBirthDate - MinBirthDate) < Age_uncert)%>%
    # Find Depart ages of all inds in adults table:
    mutate(ageAds = ceiling(as.numeric(DepartDate - BirthDate) / 365.25))%>%
    #increase the table to have a row per individual and per age
    tidyr::uncount(weights = ageAds, .id = "n", .remove = F) %>%
    mutate(Age = ageAds - n,
           Date_age = BirthDate + lubridate::years(Age))%>%
    distinct()
  
  if(Global){
    ADULTS <- ADULTS%>%
      #paste collection for each individual and age
      left_join(collection%>%as_tibble%>%
                  select(AnimalID, ScopeType, ChangeDate)%>%rename(anonID = AnimalID)%>% distinct(), 
                by = "anonID", relationship = "many-to-many")%>%
      mutate(dist = purrr::map2_dbl(Date_age, ChangeDate, difftime)) %>%
      filter(dist >= 0)%>%
      group_by(anonID, Age) %>%
      mutate(maxtime = min(dist)) %>%
      ungroup()%>%
      filter(maxtime  == dist) %>%
      dplyr::select(-maxtime , -dist, -ChangeDate, -n)%>%
      filter(ScopeType == "Global")
    
    parent <- parent %>%
      filter(ParentCollectionScopeType == "Global",
             OffspringCollectionScopeType == "Global")
  }
  
  # Number of adult records:
  fertSumm$Nadults <- length(unique(ADULTS$anonID))
  
  
  if ( fertSumm$Nadults > 0) {
    
    #select parents and remove duplicated rows
    subpar <- parent %>%
      filter(ParentAnonID %in% unique(ADULTS$anonID))%>% 
      left_join(ADULTS%>%select(anonID, BirthDate,DepartDate, birthType)%>%distinct(),
                relationship = "many-to-many",
                by =c("ParentAnonID" = "anonID"))%>%
      rename(Parent_BirthDate = BirthDate,
             Parent_DepartDate = DepartDate,
             Parent_birthType = birthType)%>% 
      select(-c("ParentType", "ParentTypeID"))%>%
      distinct() %>% 
      #Offsping not in coresubset
      filter(AnonID %in% offspSub$anonID)%>%
      left_join(offspSub%>%select(anonID, BirthDate, Sex, firstInst),
                by =c("AnonID" = "anonID"))%>%
      rename(Offspring_BirthDate = BirthDate,
             Offspring_Inst = firstInst)%>%
      mutate(ageBirth = as.numeric(Offspring_BirthDate - 
                                     Parent_BirthDate) / 365.25)%>%
      left_join(moves%>%as_tibble%>%
                  select(AnonID, To, Date), 
                by = c("ParentAnonID" = "AnonID"), relationship = "many-to-many")%>%
      mutate(dist = purrr::map2_dbl(Offspring_BirthDate , Date, difftime)) %>%
      filter(dist >= 0)%>%
      group_by(AnonID) %>%
      mutate(mintime = min(dist)) %>%
      ungroup()%>%
      filter(mintime  == dist) %>%
      dplyr::select(-mintime , -dist, -Date)%>%
      rename(currentInst = To)%>%
      distinct()
    
    
    # Number of births
    fertSumm$NOffsp <- length(unique(subpar$AnonID))
    fertSumm$NParent <- length(unique(subpar$ParentAnonID))
    
    
    
    if (fertSumm$NOffsp   > 0 ) {
      
      
      #Keep only repro with age of parents at birth >0 
      subpar <- subpar%>%
        ## VERIFIER QUE J4AI Qu'une ligne par couple parent/enfantXXXXXXXXXXXXXXXXXXXXXXX
        group_by(AnonID, ParentOriginType, ParentAnonID,  
                 OffspringCollectionScopeType, ParentCollectionScopeType, 
                 Parent_BirthDate, Parent_birthType, Offspring_BirthDate,
                 Sex, ageBirth, currentInst,  Offspring_Inst)%>%
        summarise(Probability = max(Probability))%>%
        filter(ageBirth>0)%>%ungroup()
      
      fertSumm$NOffsp_age <- length(unique(subpar$AnonID))
      fertSumm$NParent_age <- length(unique(subpar$ParentAnonID))
      
    }
  }
  
  # XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
  # # Find which individuals are in the contraception data:
  #           ADULTS <- ADULTS %>% 
  #             left_join(newContra%>%as_tibble%>%mutate(anonID = as.numeric(anonID)), 
  #                       by = "anonID", relationship = "many-to-many")%>%
  #             mutate(date = case_when(
  #               is.na(date)~ as_date(Date_age), 
  #               .default =as_date(date)))%>%
  #             mutate(dist = purrr::map2_dbl(Date_age , date, difftime)) %>%
  #             filter(dist >= 0)%>%
  #             group_by(anonID) %>%
  #             mutate(mintime = min(dist)) %>%
  #             ungroup()%>%
  #             filter(mintime  == dist) %>%
  #             mutate(contra = case_when(
  #               is.na(status) ~ FALSE, 
  #               status =="Inactive" ~ FALSE,
  #               .default = TRUE))%>%
  #             dplyr::select(-mintime , -dist, -date, -method, -status)%>%
  #             distinct()
  #           
  
  
  return(list(Adults = ADULTS, Reprodata = subpar, summary =  fertSumm))
  
  
  
}
```

```{r examples-Rep_prepdata}
data(core)
data(collection)
data(parent)
data(moves)

Data <- Rep_prepdata (coresubset = core, collection, parent, moves,
                      BirthType_parent = "Captive", BirthType_offspring = "Captive"
)
```

```{r tests-Rep_prepdata}
test_that("Rep_prepdata works", {
  data(core)
  data(collection)
  data(parent)
  data(moves)
  
  Data <- Rep_prepdata (coresubset = core, collection, parent, moves,
                        BirthType_parent = "Captive", BirthType_offspring = "Captive"
  )
  expect_named(Data, c("Adults", "Reprodata", "summary"))
  expect_named(Data$summary, c("Nbirths", "Nadults", "NOffsp", "NParent", "NOffsp_age", "NParent_age"))
  expect_true(is.numeric(Data$summary%>%as.numeric()))
  expect_named(Data$Adults, c(names(core), "ageAds", "Age", "Date_age", "ScopeType"))
  expect_named(Data$Reprodata, c("AnonID", "ParentOriginType", 
                                 "ParentAnonID", "OffspringCollectionScopeType", 
                                 "ParentCollectionScopeType", "Parent_BirthDate", 
                                 "Parent_birthType", "Offspring_BirthDate", "Sex",
                                 "ageBirth", "currentInst", "Offspring_Inst", 
                                 "Probability"))
})
```


# Seasonnality

```{r function-Rep_season}
#' my_fun Title
#'
#' @return 1
#' @export
#'
#' @examples
Rep_season <- function() {
  # --------------------------- #
  # Seasonality in births:
  # --------------------------- #
  # Extract birth dates, years and months:
  if ("All" %in% sexCats) {
    if (sexDat$summary$All$N > 0) {
      seasDat <- sexDat$data$All
      nSeas <- nrow(seasDat)      
    } else {
      nSeas <- 0
    }
  } else if (sexDat$summary$Female$N > 0 & sexDat$summary$Male$N > 0) {
    seasDat <- rbind(sexDat$data$Female, sexDat$data$Male)
    nSeas <- nrow(seasDat)
  } else if (sexDat$summary$Female$N > 0) {
    seasDat <- sexDat$data$Female
    nSeas <- nrow(seasDat)
  } else if (sexDat$summary$Male$N > 0) {
    seasDat <- sexDat$data$Male
    nSeas <- nrow(seasDat)
  } else {
    nSeas <- 0
  }
  
  
  # Hemispheres:
  seasNames <- c("All", "North", "South")
  
  
  # Fill up no analysis case:
  for (ii in 1:3) {
    seas[[seasNames[ii]]] <- list(analyzed = FALSE, prob = NULL, Nbirth_raw = nSeas, Nbirth = 0, 
                                  plv = list(xysc = NULL, levs = NULL, plmarg =  c(0, 1)), error = "", Nerr = 0, L80 = NULL)
  }
  
  # Available data:
  if (nSeas > 0) {
    # Month indices and labels:
    mon <- 1:12
    monName <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", 
                 "Sep", "Oct", "Nov", "Dec")
    
    # Extract birth dates, years and months:
    DATAsea <- seasDat%>%
      filter(MinBirthDate == MaxBirthDate,
             grepl("Captive", seasDat$birthType),
             FirstCollectionScopeType == "Global")
    
    if (nrow(DATAsea) > 0) {
      
      DATAsea <- DATAsea%>%
        mutate(birthYear = year(BirthDate),
               birthDay <-(julian(as.Date(sprintf("%s-12-31", birthYear), 
                                          format = "%Y-%m-%d")) - 
                             julian(BirthDate)) / 365 * 2 * pi,
               birthMonth = month(BirthDate))
      
      # Find which are from the northern and southern hemisphere:
      spInsts <- subinst[DATAsea$firstInst, ]
      idnorth <- which(spInsts$Latitude > 0)
      idsouth <- which(spInsts$Latitude < 0)
      
      # Store in report list:
      for (ii in 1:3) {
        DATAseason <- DATAsea
        # Find which are from the northern and southern hemisphere:
        if (ii == 2) {
          DATAseason <- DATAsea[idnorth,]
        } 
        if (ii == 3) {
          DATAseason <- DATAsea[idsouth,]
        }
        seas[[seasNames[ii]]]$Nbirth <-nrow(DATAseason)
        
        if(length(unique(DATAseason$firstInst))>1){
          if (nrow(DATAseason) >= minNseas) {
            
            #Find 80% of the births
            L_80 = quantile(yday(DATAseason$BirthDate), c(0.1,0.9))
            seas[[seasNames[ii]]]$L80 =as.numeric(L_80[2]-L_80[1])
            # Find number of births per month:
            moncounts <- sapply(mon, function(mm) {
              length(which(DATAseason$birthMonth == mm))
            })
            
            # Proportion of births per month:
            monprob <- moncounts / sum(moncounts)
            names(monprob) <- monName
            
            # Plot variables:
            xysc <- ceiling(max(monprob) * 20) / 20 + 0.05
            levs <- seq(0.05, xysc - 0.05, 0.05)
            plmarg <- c(-1.01, 1.01) * xysc
            seas[[seasNames[ii]]]$plv <- list(xysc = xysc, levs = levs, plmarg = plmarg)
            
            # Store outputs:
            seas[[seasNames[ii]]]$analyzed = TRUE
            seas[[seasNames[ii]]]$prob = monprob
            
            
          }else{
            seas[[seasNames[ii]]]$error = "Nbirth <= minNseas"
            seas[[seasNames[ii]]]$Nerr = 4}
        }else{
          seas[[seasNames[ii]]]$error = "Data from 1 Institution"
          seas[[seasNames[ii]]]$Nerr = 3}
      }
      
    }else{ 
      for (ii in 1:3) {
        seas[[seasNames[ii]]]$error = "No exact birth month"
        seas[[seasNames[ii]]]$Nerr = 2
      }
    }
    
  }else{     
    for (ii in 1:3) {
      seas[[seasNames[ii]]]$error = "NThres == 0"
      seas[[seasNames[ii]]]$Nerr = 1
    }
  }
  
  
}
```

```{r examples-Rep_season}

```

```{r tests-Rep_season}
test_that("Rep_season works", {
  
})
```


# my_fun

```{r function-my_fun}
#' my_fun Title
#'
#' @return 1
#' @export
#'
#' @examples
my_fun <- function() {
  1
}
```

```{r examples-my_fun}
my_fun()
```

```{r tests-my_fun}
test_that("my_fun works", {
  
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_repro.Rmd", vignette_name = "Reproduction")
```

