---
title: "Load data and run taxon profiles"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Run Taxon Profiles

`run_txprofile()` loads the data needed and runs or updates demographic analyses for the list of species selected. It uses the following main arguments (see the help of the function for more):

* `taxa` 
* `Species_list
* `ZIMSdir` directory where to find data
* `AnalysisDir`  directory where to save results
* `PlotDir`: Directory to save the plots of the best models
* `extractDate `  Date of data extraction
* `minDate`  Earlier date to include data
* `Sections` names of the sections to update in the taxon profile results: "sur", "rep" and/or "gro"
* `sexCats` Male, Female or/and All
* `models_sur` names of the survival basta models to run and compared: "G0", "EX", "LO" and/or "WE".
* `models_gro` indicating the growth models that need to be run and compared. The following models are supported : logistic, gompertz, chapmanRichards, vonBertalanffy, polynomial.

The result files are saved using the following directory \code{{analysisDir}\{taxa}\{species}.Rdata}, where `analysisDir` and `taxa` are argument of the function. `species` is the latin name of each species



```{r function-my_fun}
#' Run Taxon Profiles
#' 
#' Load data needed and run or update demographic analyses for the list of species selected
#'
#' @param taxa  \code{character} the name of the taxa studied
#' @param Species_list \code{vector of character} Species selected
#' @param ZIMSdir \code{character} directory where to find data
#' @param AnalysisDir \code{character} directory where to save results
#' @param PlotDir \code{character} Directory to save the plots. Default = NULL, no plot is saved
#' @param extractDate \code{character 'YYYY-MM-DD'} Date of data extraction
#' @param minDate \code{character 'YYYY-MM-DD'} Earlier date to include data
#' @param Sections \code{vector of character} names of the sections to update in the taxon profile results: "sur", "rep" and/or "gro". Default = c("sur", "rep", "gro")
#' @param erase_previous \code{logical} whether the current result file should be deleted (before being replaced). Default = FALSE
#' @param sexCats \code{character} Male, Female or All Default =  "All"
#' @param inparallel \code{logical} Whether this function is run in parallel on different computer. In other words: should the species list be divided? Default = FALSE
#' @param ipara \code{numeric} Id number of the computer used to select the species to run
#' @param npara \code{numeric} number of computers used in parallel
#' @param minN \code{numeric} Minimum number of individuals. Default = 50
#' @param maxOutl \code{numeric} Maximum threshold for the longevity distribution. Default = NULL
#' @param spOutLev \code{vector of character} List of species for which the Maximum threshold for the longevity distribution.XXXXXXXXXXXXXXXXXXXXXXXXX
#' @param minNsur \code{numeric} Minimum number of individuals to run the survival analysis. Default = 50
#' @param minlx  \code{numeric} between 0 and 1. Minimum reached survivorship from the raw Kaplan Meier analysis needed to run the survival analysis. Default = 0.1
#' @param MinBirthKnown  \code{numeric} between 0 and 1. Minimum proportion of individuals with a known birth month in the data. Default = 0.3
#' @param models_sur \code{vector of characters} names of the survival basta models to run: "G0", "EX", "LO" and/or "WE". see ?basta for more information. Default = "GO"
#' @param shape \code{character} shape of the survival basta model to run: "simple", "Makeham", "bathtub".  see ?basta for more information. Default = "simple"
#' @param niter  \code{numeric}. number of MCMC iterations to run the survival model. see ?basta for more information. Default = 25000
#' @param burnin  \code{numeric} Number of iterations removed so that the survival model has time to converge. see ?basta for more information. Default = 5001
#' @param thinning  \code{numeric} Number of iteration to run the survival model before saving a set of parameters. see ?basta for more information. Default = 20
#' @param nchain  \code{numeric} Number of chains to run the survival model. Default = 5001
#' @param ncpus  \code{numeric} Number of computer core to use. Default = 2
#' @param parentProb \code{numeric} XXXXXXXXXXX
#' @param minNrepro  \code{numeric} XXXXXXXXXXX
#' @param minNparepro  \code{numeric} XXXXXXXXXXX
#' @param minNseas \code{numeric} XXXXXXXXXXX
#' @param minNgro \code{numeric} Minimum number of weight needed to fit the growth models
#' @param minNIgro \code{numeric} Minimum number of unique individuals needed to fit the growth models
#' @param MeasureType \code{vector of characters} Name of the type of measurements that should be included.  Default = NULL, all measurement type are included.
#' @param models_gro \code{vector of characters} indicating the growth models that need to be fit.The following models are supported : logistic, gompertz, chapmanRichards, vonBertalanffy, polynomial. default = "vonBertalanffy"
#'
#' @return Save and replace the result file for each specie sin the list. The file is saved in {analysisDir}\\{taxa}\\{species}.Rdata
#' @export
#'
#' @examples
run_txprofile <- function(taxa, Species_list, ZIMSdir, 
                          AnalysisDir, PlotDir,
                          extractDate, minDate = "1980-01-01",
                          Sections, erase_previous = FALSE,
                          sexCats = c('Male', 'Female'), 
                          inparallel = FALSE, ipara = 1, npara = 1, 
                          minN= 50,  maxOutl =99,  spOutLev = NULL,
                          minNsur = 50, minlx = 0.1, MinBirthKnown = 0.3, 
                          models_sur = "GO", shape = "bathub",
                          niter = 25000, burnin = 5001, thinning = 20, 
                          nchain = 3, ncpus = 2,
                          parentProb = 80, minNrepro = 100, 
                          minNparepro = 30, minNseas = 50, 
                          minNgro =100,minNIgro = 50, MeasureType = "",
                          models_gro = "vonBertalanffy"
                          
){
  
  assert_that(is.character(taxa))
  assert_that(taxa %in% c("Mammalia", "Aves", "Reptilia", "Amphibia", 
                          "Chondrichthyes", "Actinopterygii"),
              msg = "taxa must one of 'Mammalia', 'Aves', 'Reptilia', 'Amphibia', 
                          'Chondrichthyes', or 'Actinopterygii'")
  assert_that(is.character(Sections))
  assert_that(all(Sections %in% c("sur", "gro", "rep")))
  assert_that(is.character(Species_list))
  minDate = lubridate::as_date(minDate)
  extractDate = lubridate::as_date(extractDate)
  assert_that(is.logical( erase_previous))
  assert_that(is.logical(inparallel))
  assert_that(is.numeric( iparal))
  assert_that(is.numeric( npara))
  assert_that(ipara <= npara)
  assert_that(is.numeric( maxOutl))
  assert_that( maxOutl <= 100)
  assert_that(is.numeric(minNsur))
  assert_that(minNsur > 0)
  assert_that(is.numeric(minlx))
  assert_that(minlx > 0)
  assert_that(minlx < 1)
  assert_that(is.numeric(MinBirthKnown))
  assert_that(MinBirthKnown > 0)
  assert_that(MinBirthKnown <1)
  assert_that(is.numeric(niter))
  assert_that(niter > 0)
  assert_that(is.numeric(burnin))
  assert_that(burnin > 0)
  assert_that(burnin < niter)
  assert_that(is.numeric(thinning))
  assert_that(thinning > 0)
  assert_that(thinning < niter)
  assert_that(is.numeric(nchain))
  assert_that(nchain > 0)
  assert_that(is.numeric(ncpus))
  assert_that(ncpus > 0)
  assert_that(is.character(models_sur))
  assert_that(all(models_sur %in% c("GO", "EX", "LO", "WE")))
  assert_that(is.character(shape))
  assert_that(all(shape %in% c("simple", "bathtub", "Makeham")))
  assert_directory_exists(ZIMSdir)
  assert_directory_exists(AnalysisDir)
   assert_directory_exists(PlotDir)
 
  assert_that(is.numeric(parentProb))
  assert_that(parentProb > 0)
  assert_that(is.numeric(minNrepro))
  assert_that(minNrepro > 0)
  assert_that(is.numeric(minNparepro))
  assert_that(minNparepro > 0)
  assert_that(is.numeric( minNseas))
  assert_that( minNseas > 0)
  
  assert_that(is.numeric( minNIgro))
  assert_that( minNIgro > 0)
  assert_that(is.numeric(minNgro))
  assert_that(minNgro > 0)
  assert_that(is.character(MeasureType))
  assert_that(is.character(models_gro ))
  assert_that(all(models_gro %in% c("logistic", "gompertz", "chapmanRichards", "vonBertalanffy", "polynomial", "gam")), msg = "The growth models supported are: logistic, gompertz, chapmanRichards, vonBertalanffy, and polynomial, in addition to GAM.")
  
  assert_that(is.character( sexCats))
   assert_that(all(sexCats %in% c("Male", "Female", "All")))
  
  
  
  # ======================#
  # ==== Load data ========
  # ======================#
  tables = c( "core", "collections")
  if("gro" %in% Sections){
    tables = c(tables, "weights")
  }
  if("rep" %in% Sections){
    tables = c(tables, "parents", "contraception")
  }
  Load_Zimsdata	(taxa = taxa, ZIMSdir = ZIMSdir, 
                 extractDate, 
                 type = tables,
                 silent = TRUE) 
  
  # ======================#
  # ==== Species List =====
  # ======================#
  if( Species_list == "All"){
    spAll <- unique(core$binSpecies)
  }
  if (inparallel){
    # IDs of species to run per version:
    idsprun <- seq(ipara, length(spAll), by = npara)
  }else{
    idsprun <- c(1:length(spAll))}
  
  # ======================#
  # ==== RUN ANALYSES: ====
  # ======================#
  # Start counter:
  icount <- 0
  
  # Loop over species
  for (isp in idsprun) {
    # Extract species:
    species <- spAll[isp]
    
    # # progress count:
    icount <- icount + 1
    
    #subset Tables
    if("gro" %in% Sections){
      weight_spe <- weights%>%
        filter(binSpecies == "species")
    }else{
      weight_spe <-NULL
    }
    if("rep" %in% Sections){
      contra_spe <- contraceptions%>%
        filter(binSpecies == "species")
      parent_spe <- parents()%>%
        filter(binSpecies == "species")
    }else{
      parent_spe <- contra_spe <-NULL
    }
    
    # Report progress:
    cat("\n====================\nClass:   ", taxa, "\nSpecies: ", species, 
        "\nProgress:", round(icount / length(idsprun) * 100, 0), 
        "%\n=====================\n")
    cat("Species running... ")
    
    #Load previous taxon profile to update it
    if(erase_previous){
      Repout = list()
    }else{
      if(any( stringr::str_detect(list.files("ISRdata/global/rdata/", species)))){
        load(glue::glue("ISRdata/global/rdata/{species}.RData"))
      }else{Repout = list()}
    }
    maxOutl= NULL
    if (species %in% spOutLev) maxOutl <- 99.9
    
    
    # Create report:
    repout <- tx_report(species, taxa, 
                        coresubset, collection, 
                        PlotDir = PlotDir, extractDate= extractDate,
                        weights = weight_spe, parents = parent_spe, contraceptions = contra_spe,
                        repout = repout, Sections = Sections,
                        sexCats = sexCats, minN= minN, minDate= minDate, 
                        maxOutl = maxOutl, minlx = minlx, 
                        minNsur = minNsur, MinBirthKnown = MinBirthKnown,
                        models_sur = models_sur, shape = shape,
                        niter = niter, burnin =  burnin, thinning = thinning,
                        nchain = nchain,  ncpus = ncpus,
                        parentProb = parentProb, minNrepro = minNrepro,
                        minNparepro = minNparepro, minNseas = minNseas, 
                        minNgro =minNgro,minNIgro = minNIgro, 
                        MeasureType = MeasureType, models_gro = models_gro
    )
    
    # ----------------------- #
    # ---- Save results: ----
    # ----------------------- #
    # Save species list:
    save("repout", file = glue::glue("{analysisDir}/{species}.RData"))
    cat(" done.\n")
  }
}

```

```{r examples-run_txprofile}
# # Here is an example but use directly your own ZIMSdir(directory to find data) and analysisDir(directory to save analysis)
# file = system.file("coretest.csv", package = 'ISRverse')
# ZIMSdir = dirname(file)
# AnalysisDir = paste0(tempdir(check = TRUE),'\\temp')
# PlotDir = paste0(tempdir(check = TRUE),'\\temp\\plot')
# dir.create(AnalysisDir)
# dir.create(PlotDir)
# 
# #This code run the survival analysis for gorilla
# out <- run_txprofile(taxa = "Mammalia", Species_list = c('Gorilla gorilla'), 
#                           ZIMSdir = RastDir, AnalysisDir = AnalysisDir,
#                           PlotDir = PlotDir,
#                           extractDate = "", 
#                           minDate = "1980-01-01",
#                           Sections = "sur", 
#                           sexCats = c('Male', 'Female'),
#                          niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3
# )
# 
# list.files(PlotDir)
# list.files(AnalysisDir)
# 
# unlink(AnalysisDir, recursive = TRUE)

```


This function can be run on parallel on several computer on the same species list. In this case the following arguments need to be used:
* `inparallel` Whether this function is run in parallel on different computer. In other words: should the species list be divided? Default = FALSE
* `ipara` Id number of the computer used to select the species to run
* `npara` number of computers used in parallel
.

```{r examples2-run_txprofile}
#' \dontrun{
#' # Here is an example but use directly your own ZIMSdir(directory to find data) and analysisDir(directory #' to save analysis)
#' file = system.file("coretest.csv", package = 'ISRverse')
#' ZIMSdir = dirname(file)
#' analysisDir = paste0(tempdir(check = TRUE),'\\temp')
#' PlotDir = paste0(tempdir(check = TRUE),'\\temp\\plot')
#' dir.create(analysisDir)
#' dir.create(PlotDir)
#' 
#' #Here I want to split my analysis on two computers
#' #Run on computer 1:
#' out <- run_txprofile(taxa = "Mammalia", Species_list = c('Gorilla gorilla', "Capreolus capreolus"), 
#'                           ZIMSdir = RastDir, analysisDir = analysisDir,
#'                          PlotDir = PlotDir,
#'                        extractDate = "", 
#'                           minDate = "1980-01-01",
#'                           Sections = "sur", 
#'                           sexCats = c('Male', 'Female'), 
#'                           inparallel = TRUE, 
#'                           ipara = 1, npara = 2
#' )
#' 
#' 
#' #Run on computer 2:
#' out <- run_txprofile(taxa = "Mammalia", Species_list = c('Gorilla gorilla', "Capreolus capreolus"), 
#'                           ZIMSdir = RastDir, analysisDir = analysisDir,
#'                           PlotDir = PlotDir,
#'                       extractDate = "", 
#'                           minDate = "1980-01-01",
#'                           Sections = "sur", 
#'                           sexCats = c('Male', 'Female'), 
#'                           inparallel = TRUE,
#'                            ipara = 2, npara = 2
#' )
#' list.files(PlotDir)
#' list.files(AnalysisDir)
#'
#' unlink(analysisDir, recursive = TRUE)
#' }
```

```{r tests-run_txprofile}
# file = system.file("coretest.csv", package = 'ISRverse')
# ZIMSdir = dirname(file)
# analysisDir = paste0(tempdir(check = TRUE),'\\temp')
# PlotDir = paste0(tempdir(check = TRUE),'\\temp\\plot')
# dir.create(analysisDir)
# dir.create(PlotDir)
# 
# test_that("run_txprofile works", {
#  
# out <- run_txprofile(taxa = "Mammalia", Species_list = "All", 
#                           ZIMSdir = RastDir, analysisDir = analysisDir,
#                           PlotDir = PlotDir, extractDate = "", 
#                           minDate = "1980-01-01",
#                           Sections = "sur", 
#                           sexCats = c('Male', 'Female')
# )
# 
# expect_true(file.exists(paste(analysisDir, "Gorilla_gorilla.Rdata", sep = '\\')))
# expect_true(file.exists(paste(PlotDir, 'Gorilla_gorilla_LongThres.pdf', sep = '\\')))
# expect_true(file.exists(paste(PlotDir, 'Gorilla_gorillasur.pdf', sep = '\\')))
# expect_true(file.exists(paste(PlotDir, 'Gorilla_gorilla_outliers.png', sep = '\\')))
# expect_true(file.exists(paste(PlotDir, 'Gorilla_gorilla_growth.png', sep = '\\')))
# 
# })
# unlink(RastDirgp, recursive = TRUE)

```

# Summary tables & plots

`make_summary()` produces summary tables and plots of the demographic analyses made for all species. It uses the following arguments:

* `AnalysisDir` character} directory where to find the .Rdata files
* `SaveDir` character} directory where to save summary plots and tables
* `taxaList` names of the taxa studied.
* `BySex` list of the taxa names indicating the sexes analyzed. 
* `Sections` names of the sections to update in the taxon profile results: "sur", "rep" and/or "gro"

it returns the main summary table

```{r function-make_summary}
#' Summarize taxon profiles analyses
#' 
#' Produce summary tables and plots of the demographic analyses made for all species
#'
#' @param AnalysisDir  \code{character} directory where to find the .Rdata files
#' @param SaveDir  \code{character} directory where to save summary plots and tables
#' @param namefile \code{character} Suffix to add to the name of files produced if needed. Default = ""
#' @param taxaList \code{vector of character} names of the taxa studied. Default= "Mammalia"
#' @param BySex \code{list} of the taxa names indicating the sexes analyzed. Default=list(Mammalia = c("Male", "Female"))
#' @param Sections \code{vector of character} names of the sections to update in the taxon profile results: "sur", "rep" and/or "gro". Default = c("sur", "rep", "gro")
#'
#' @return It saves summary tables and plots for each Sections and a general summary table. It returns the main summary table
#' @export
#' 
#' @importFrom ggplot2 ggplot scale_fill_brewer facet_wrap geom_bar labs position_dodge coord_flip
#' @importFrom ggpubr ggarrange
#' 
#' @examples
make_summary <- function (AnalysisDir, SaveDir, namefile = "",
                          taxaList = "Mammalia", 
                          BySex = list(Mammalia = c("Male", "Female")) , 
                          Sections = c("sur", 'rep', 'gro')
){
  assert_that(is.character(taxaList))
  assert_that(taxaList %in% c("Mammalia", "Aves", "Reptilia", "Amphibia", 
                              "Chondrichthyes", "Actinopterygii"),
              msg = "taxa must one of 'Mammalia', 'Aves', 'Reptilia', 'Amphibia', 
                          'Chondrichthyes', or 'Actinopterygii'")
  assert_that(is.character(Sections))
  assert_that(all(Sections %in% c("sur", "gro", "rep")))
  checkmate::assert_directory_exists(AnalysisDir)
   checkmate::assert_directory_exists(SaveDir)
  assert_that(is.character(namefile))
  assert_that(is.list(BySex))
  assert_that(taxaList %in% names(BySex), msg = "BySex should be a list with names identical to taxaList")
  
  # List of available SRGs:
  SRGlist <- list.files(AnalysisDir, pattern = ".Rdata")
  assert_that(length(SRGlist) > 0, 
              msg = glue::glue("There are no result file in {analysisDir}"))
  SRGsps <- gsub(".RData", "", SRGlist)
  
  # Start counter:
  icount <- 0
  for (taxa in taxaList) {
    sexCats <- BySex [[taxa]]
    
    icount <- icount + 1
    
    table<-tibble(Class = rep(taxa, length(SRGlist)*length(sexCats)),
                  Species = rep(SRGsps, each = length(sexCats)),
                  Sex = rep(sexCats, length(SRGlist)),
                  Nraw = numeric(1),
                  Ndate = numeric(1),
                  Nglobal = numeric(1),
                  Nalive = numeric(1),
                  firstDate = date(),
                  maxAgeraw = numeric(1),
                  extractdate = date(),
                  GapThresh = numeric(1),
                  NThres = numeric(1)
    )
    
    # Taxa data table:
    if ("surv" %in% Sections){
      tempsur <- table%>%
        mutate(NBasta = numeric(1), 
               Ndead = 0, 
               lxMin = numeric(1),
               maxAlive = numeric(1),
               outLev = numeric(1),
               analyzed = logical(1), 
               Nerr = numeric(1), 
               error = character(1))
    }else{tempsur = tibble()}
    
    if ("gro" %in% Sections){
      tempgro <- table%>%
        mutate(NWeight_raw = numeric(1), NInd_raw = numeric(1), 
               NWeight_val = numeric(1), NInd_val = numeric(1), 
               NWeight_age = numeric(1), NInd_age = numeric(1), 
               agemat = numeric(1), 
               NJuv = numeric(1),NJuv_keep = numeric(1),
               NAd = numeric(1), NAd_keep = numeric(1),
               NWeight = numeric(1), NInd = numeric(1), 
               analyzed = logical(1), 
               Nerr = numeric(1), error = character(1))
    }else{tempgro = tibble()}
    if ("rep" %in% Sections){
      temprep <- table%>%
        mutate()
    }else{temprep = tibble()}
    
    # Loop over species
    for (isp in SRGsps) {
      # SRG file:
      load(glue::glue("{ResultFileDir}/{isp}.RData"))
      
      table = table%>%
        rows_update(repout$General %>% mutate(Species = isp, Sex = "All"), 
                    by = c("Species", "Sex"), unmatched = "ignore")
      
      # Fill up data list:
      for (sx in sexCats) {
        
        table = table%>%
          rows_update(repout$summar[[sx]]%>%as_tibble() %>% 
                        mutate(Species = isp, Sex = sx), 
                      by = c("Sex", "Species"), unmatched = "ignore")
        
        if("sur" %in% Sections){
          tempsur <- tempsur%>%
            rows_update(repout$surv[[sx]]$summary%>%as_tibble() %>% 
                          mutate(Species = isp, Sex = sx), 
                        by = c("Sex", "Species"), unmatched = "ignore")
        }
        
        if("gro" %in% Sections){
          tempgro <- tempgro%>%
            rows_update(repout$weig[[sx]]$summary%>%as_tibble() %>% 
                          mutate(Species = isp, Sex = sx), 
                        by = c("Sex", "Species"), unmatched = "ignore")
        }
        
        if("rep" %in% Sections){
          temprep <- temprep%>%
            rows_update(repout$rep[[sx]]$summary%>%as_tibble() %>% 
                          mutate(Species = isp, Sex = sx), 
                        by = c("Sex", "Species"), unmatched = "ignore")
        }
      }
    }
    
    if("sur" %in% Sections){
      tempsur <- tempsur%>%
        rows_update(table, 
                    by = c("Sex", "Species"), unmatched = "ignore")
    }
    
    if("gro" %in% Sections){
      tempgro <- tempgro%>%
        rows_update(table, 
                    by = c("Sex", "Species"), unmatched = "ignore")
    }
    
    if("rep" %in% Sections){
      temprep <- temprep%>%
        rows_update(table, 
                    by = c("Sex", "Species"), unmatched = "ignore")
    }
    
    if (icount == 1) {
      SummTab <-table
      SurTab <- tempsur
      RepTab <- temprep
      GroTab <- tempgro
    } else {
      SummTab <- rbind(SummTab, table)
      SurTab <- rbind(SurTab, tempsur)
      RepTab <- rbind(RepTab, temprep)
      GroTab <- rbind(GroTab, tempgro)
    }
    
  }
  
  if("sur" %in% Sections){  
    SummTab <-SummTab%>%
      left_join(SurTab%>%
                  select(c(Species, Sex, NBasta, Analyzed, error))%>%
                  rename(Surv_Ana = Analyzed, Surv_error = error), 
                by = c("Species", "Sex"))
    
    utils::write.csv(SurTab, file = glue::glue("{savedir}/SRGs_Survival{namefile}.csv"),
              row.names = FALSE)
    
    Surtabsum <- SurTab %>% 
      mutate(error =ifelse(error =="", "Analyzed",error),
             error = factor(error, levels = c('Analyzed','NThres == 0', 
                                              "Nglobal = 0", "%known births < 0.3",
                                              "Data from 1 Institution",
                                              "Nbasta < minNsur", "lxMin > minlx",
                                              "no DIC from Basta", "lx[MLE]<0.1",
                                              "Min(Life_exp)>2"), 
                            ordered = T)) %>% 
      group_by(Class, Sex, error)%>% summarize(N = n())
    p<- ggplot(data = Surtabsum, aes(x = error, y = N, fill = Sex)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      scale_fill_brewer(palette = "Spectral")+
      labs(x = "") + facet_wrap(~ Class, scales = "free")+
      coord_flip()
    ggsave( glue::glue("{savedir}/Survival_error{namefile}.pdf"), p, width = 20, height = 6)
    
  }
  
  if("rep" %in% Sections){ 
    SummTab <-SummTab%>%
      left_join(RepTab%>%
                  select(Species, Sex, NOffsp_raw, NParent_raw, NOffsp,NParent, 
                         NAdult_rep, Fert_Analyzed,Fert_error,
                         NOffsp_prob,NParent_prob, NReprEvent,Litt_Analyzed,Litt_error,
                         SeasNorth_Analyzed, SeasNorth_error, SeasNorth_Nbirth,
                         SeasSouth_Analyzed, SeasSouth_error, SeasSouth_Nbirth)%>%
                  rename(Fert_Ana = Fert_Analyzed,
                         Litt_Ana = Litt_Analyzed,
                         SeasNorth_Ana = SeasNorth_Analyzed,
                         SeasSouth_Ana = SeasSouth_Analyzed), 
                by = c("Species", "Sex"))
    
    
    utils::write.csv(RepTab, file =  glue::glue("{savedir}/SRGs_Reproduction{namefile}.csv", globDir),
              row.names = FALSE)
    
    Ferttabsum <- RepTab  %>% tidyr::drop_na(Fert_error)%>% 
      mutate(Fert_error = ifelse(Fert_error =="", "Analyzed",Fert_error),
             Fert_error = factor(Fert_error, levels =c('Analyzed','NThres == 0', 
                                                       "NAdult == 0", "NOffspr_age == 0",
                                                       "NParent_bd == 0", 
                                                       "Data from 1 Institution",
                                                       "NOffsp < minNrepro",
                                                       "NParent < minNparepro"), 
                                 ordered = T)) %>% 
      group_by(Class, Sex, Fert_error)%>% summarize(N = n())
    fert<- ggplot(data=Ferttabsum, aes(x=Fert_error, y=N, fill=Sex)) +
      geom_bar(stat="identity", position=position_dodge()) +
      scale_fill_brewer(palette="Spectral")+
      labs(x = "Fertility")+
      facet_wrap(~Class, nrow = 1, scales = "free")+
      coord_flip()
    Litttabsum <- RepTab  %>% tidyr::drop_na(Litt_error) %>% 
      filter(Sex!="Male")%>%
      mutate(Litt_error =ifelse(Litt_error =="", "Analyzed",Litt_error),
             Litt_error =ifelse(Litt_error =="NOffsp  < minNrepro",
                                "NOffsp < minNrepro",Litt_error),
             Litt_error = factor(Litt_error, levels = c('Analyzed','NThres == 0', 
                                                        "NParent_bd == 0", 
                                                        "NAdult == 0", "NOffspr_age == 0",
                                                        "NOffsp < minNrepro", 
                                                        "NParent < minNparepro",
                                                        "Data from 1 Institution",
                                                        "NOffsp_prob < minNrepro"), 
                                 ordered = T))%>% 
      group_by(Class, Sex, Litt_error)%>% summarize(N = n())
    lit<- ggplot(data=Litttabsum, aes(x=Litt_error, y=N, fill=Sex)) +
      geom_bar(stat="identity", position=position_dodge()) +
      scale_fill_brewer(palette="Spectral")+
      labs(x = "Litter Size")+
      facet_wrap(~Class, nrow = 1, scales = "free")+
      coord_flip()
    SeaNtabsum <- RepTab  %>% tidyr::drop_na(SeasNorth_error)%>%
      filter(Sex=="Female" | Class == "Actinopterygii")%>%
      mutate(SeasNorth_error =ifelse(SeasNorth_error =="", "Analyzed",SeasNorth_error),
             SeasNorth_error = factor(SeasNorth_error, levels = c('Analyzed',"NThres == 0", 
                                                                  "No exact birth month",
                                                                  "Data from 1 Institution",
                                                                  "Nbirth <= minNseas"), 
                                      ordered = T)) %>% 
      group_by(Class, SeasNorth_error)%>% summarize(N = n())
    seaN<- ggplot(data=SeaNtabsum, aes(x=SeasNorth_error, y=N)) +
      geom_bar(stat="identity", position=position_dodge()) +
      labs(x = "Seasonality North")+
      facet_wrap(~Class, nrow = 1, scales = "free")+
      coord_flip()
    SeaStabsum <- RepTab  %>% tidyr::drop_na(SeasSouth_error)%>% 
      filter(Sex=="Female" | Class == "Actinopterygii")%>%
      mutate(SeasSouth_error =ifelse(SeasSouth_error =="", "Analyzed",SeasSouth_error),
             SeasSouth_error = factor(SeasSouth_error, levels = c('Analyzed',"NThres == 0", 
                                                                  "No exact birth month", 
                                                                  "Data from 1 Institution",
                                                                  "Nbirth <= minNseas"), 
                                      ordered = T)) %>% 
      group_by(Class, SeasSouth_error)%>% summarize(N = n())
    seaS<- ggplot(data=SeaStabsum, aes(x=SeasSouth_error, y=N)) +
      geom_bar(stat="identity", position=position_dodge()) +
      labs(x = "Seasonality South")+
      facet_wrap(~Class, nrow = 1, scales = "free")+
      coord_flip()
    figure <- ggarrange(fert, lit, seaN,seaS,
                        # labels = c("A", "B", "C"),
                        ncol = 1, nrow = 4)
    ggsave( glue::glue("{savedir}/Reproduction_error{namefile}.pdf"), figure, width = 25, height = 15)
    
  }          
  
  if("gro" %in% Sections){            
    SummTab <-SummTab%>%
      left_join(GroTab%>%
                  select(Species, Sex, NWeight_raw , NWeight, Analyzed,error)%>%
                  rename(Gro_Ana = Analyzed, Gro_error = error), 
                by = c("Species", "Sex"))
    
    
    utils::write.csv(GroTab, file = glue::glue("{savedir}/SRGs_Growth{namefile}.csv", globDir),
              row.names = FALSE)
    
    Grotabsum <- GroTab  %>% tidyr::drop_na(error)%>% 
      mutate(error =ifelse(error =="", "Analyzed",error),
             error = factor(error, levels = c('Analyzed',"No known Sex", 
                                              "No weight for this sex category",
                                              "No valid weight measure", "Data from 1 Institution",
                                              "NWeight < minNgro", "NInds < minNIgro", "Model did not fit"), 
                            ordered = T)) %>%
      group_by(Class, Sex, error)%>% summarize(N = n())
    p<- ggplot(data=Grotabsum, aes(x=error, y=N, fill=Sex)) +
      geom_bar(stat="identity", position=position_dodge()) +
      scale_fill_brewer(palette="Spectral")+
      labs(x = "Growth")+facet_wrap(~Class, nrow = 2, scales = "free")+
      coord_flip()
    ggsave( glue::glue("{savedir}/Growth_error{namefile}.pdf"), p, width = 20, height = 6)
    
  }
  
  
  SummTab <-SummTab%>%
    rowwise%>%
    mutate( Any_Ana = any(across(ends_with("analyzed"))),
            All_Ana = all(across(ends_with("analyzed"))))
  
  
  utils::write.csv(SummTab, file =  glue::glue("{savedir}/SRGs_Analyses{namefile}.csv", globDir),
            row.names = FALSE)
  return(SummTab)
  
}
```

```{r examples-make_summary}
# file = system.file("gorilla gorilla.Rdata", package = 'ISRverse')
# AnalysisDir  = dirname(file)
# SaveDir = paste0(tempdir(check = TRUE),'\\temp')
# dir.create(SaveDir)
# 
# SummTab <- make_summary (AnalysisDir, SaveDir,
#                           taxaList = "Mammalia", 
#                           BySex = list(Mammalia = c("Male", "Female")) , 
#                           Sections = c("sur", 'gro')
# )
# list.files(SaveDir)
# 
# 
# unlink(SaveDir, recursive = TRUE)

```

```{r tests-make_summary}
# test_that("make_summary works", {
#   file = system.file("gorilla gorilla.Rdata", package = 'ISRverse')
# AnalysisDir  = dirname(file)
# SaveDir = paste0(tempdir(check = TRUE),'\\temp')
# dir.create(SaveDir)
# 
# SummTab <- make_summary (AnalysisDir, SaveDir,
#                           taxaList = "Mammalia", 
#                           BySex = list(Mammalia = c("Male", "Female")) , 
#                           Sections = c("sur", 'gro')
# )
# expect_true(file.exists(paste(TempDir, 'Survival_error.pdf', sep = '\\')))
# expect_true(file.exists(paste(TempDir, 'SRGs_Survival.csv', sep = '\\')))
# expect_true(file.exists(paste(TempDir, 'Growth_error.pdf', sep = '\\')))
# expect_true(file.exists(paste(TempDir, 'SRGs_Growth.csv', sep = '\\')))
# expect_true(file.exists(paste(TempDir, 'SRGs_Analyses.csv', sep = '\\')))
#   # expect_named(SummTab , c()) XXXXXXXXXXXXXXXXXXXXXXXXXX
# 
# unlink(SaveDir, recursive = TRUE)
# })

```

# Load zims core data

`Load_Zimsdata()` load ZIMS data frames of the taxa selected in a list. It uses the following arguments:

* `taxa` the name of the taxa studied`
* `ZIMSdir` directory where to find data
* `extractDate` Date of data extraction
* `type` of data to load: "core", "health", "institution", "moves", "parent", "weights", "lengths", "collections", and/or "simplified_collections"

```{r function-Load_Zimsdata}
#' Load Zims data
#' 
#' Check if selected data are already in the global environment. If not, it load them.
#'
#' @param taxa  \code{character} the name of the taxa studied
#' @param ZIMSdir \code{character} directory where to find data
#' @param type \code{vector character} the type of data to load. Default = 'core' 
#' @param extractDate \code{character 'YYYY-MM-DD'} Date of data extraction
#' @param silent \code{logical} Whether information of advancement should be printed
#' 
#' @details
#' \code{type} can take the following values : "core", "health", "institution", "moves", "parent", "weights", "lengths", "collections", "simplified_collections"
#'
#' @return A list of the data frames
#' @importFrom glue glue
#' @importFrom checkmate assert_directory_exists
#' @export
#'
#' @examples
Load_Zimsdata	<- function (taxa, 
                           ZIMSdir, 
                           extractDate, 
                           type = 'core',
                           silent = FALSE) 
{ out <- list()
  assert_directory_exists(ZIMSdir)
  assert_that(is.character(taxa))
  assert_that(taxa %in% c("Mammalia", "Aves", "Reptilia", "Amphibia", 
                          "Chondrichthyes", "Actinopterygii"),
              msg = "taxa must one of 'Mammalia', 'Aves', 'Reptilia', 'Amphibia', 
                          'Chondrichthyes', or 'Actinopterygii'")
  assert_that(is.logical( silent))
  assert_that(is.character(extractDate))
  
  assert_that(is.character(type))
  assert_that(all(type %in%  c("contraception", "core", "health", 
                               "institution", "moves", "parent", "weights", 
                               "lengths", "collections", "simplified_collections")))
  
  ZIMSfilesdate <- stringr::str_subset(list.files(ZIMSdir), pattern = extractDate)
  
  idTaxa <-  stringr::str_subset(ZIMSfilesdate, taxa)
  assert_that(length(idTaxa) > 0,
              msg =glue("Folder for {extractDate}_{taxa} not found. Verify ZIMSdir path.")
  )
  for (ty in type){
    # Check if the data have already been load
    if (exists(sym(ty), envir = globalenv())) {
      b <- eval(sym(ty))
      if (unique(b$Class) == taxa) {
        if (!silent) {
          cat(glue::glue("{ty} already loaded."))
        }
        continue <- FALSE
      }else {
        continue <- TRUE
      }
    }else {
      continue <- TRUE}
    
    if (continue) {
      idfiles <- list.files(glue("{ZIMSdir}/{idTaxa}"), glue("sci_{ty}"))
      idfilesn <- list.files(glue("{ZIMSdir}/{idTaxa}"), glue("template_{ty}"))
      
      assert_that(length(idfiles )> 0,
                  msg =glue("{ty} file for{taxa} not found.")
      )
      assert_that(length(idfiles) < 2,
                  msg = glue("More than one {ty} file for {taxa}.")
      )
      if (!silent) {
        cat("Loading {ty}")
      }
      namesvar = readr::read_delim(glue::glue("{ZIMSdir}/{idTaxa}/{idfilesn}"), delim = "@")
      b <- readr::read_delim(glue::glue("{ZIMSdir}/{idTaxa}/{idfiles}"), delim = "@", col_names = colnames(namesvar))
        out[[ty]]<- b}
    if (!silent) {
      cat("Done.\n")
    }

  }
  return(out)
}
```

```{r examples-Load_Zimsdata}

# #example of an environmental ncdf file saved with the package
# file = system.file("coretest.csv", package = 'ISRverse')
# ZIMSdirtest = dirname(file)
# 
# data <- Load_Zimsdata	(taxa = "Mammalia", 
#                            ZIMSdir = ZIMSdirtest, 
#                            extractDate = "2023-12-04", 
#                            type = 'core') 
# data$core
# unlink(ZIMSdirtest, recursive = TRUE)

```

```{r tests-Load_Zimsdata}
# test_that("Load_Zimsdata works", {
#   file = system.file("coretest.csv", package = 'ISRverse')
# RastDir = dirname(file)
# 
# data <- Load_Zimsdata	 (taxa = "Mammalia", 
#                            ZIMSdir = RastDir, 
#                            extractDate = "2023-12-04", 
#                            type = c('core','weights'),
#                            silent = TRUE) 
# expect_named(data, c("core",  "weights"))
#   expect_true(is.data.frame(core))
#   expect_true(is.data.frame(weights))
#   # expect_named(data$core, c("")) XXXXXXXXXXXXXXXXXXXXXXX
#   # expect_named(data$weights, c("")) XXXXXXXXXXXXXXXXXXXXXXX
#   unlink(ZIMSdirtest, recursive = TRUE)
# 
# })

```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_main.Rmd", vignette_name = "Taxon profiles")
```


