---
title: "flat_additional.Rmd empty"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
library(assertthat)
library(tidyverse)
library(glue)
library(extractZIMS)

```

```{r development-load}
# Load already included functions if relevant
pkgload::load_all(export_all = FALSE)
```

# Taxon profile: Main survival Analysis

`Sur_main()` runs the survival analysis (see next function : `Sur_ana()`) and the remaining life expectancy analysis (see `Sur_relex()`) on each sex category. In addition to the arguments of `Sur_ana()`, `Sur_main()` includes the following arguments:

* `data.core`the core data of the species from ZIMS including at least the following columns *anonID*, *Birth.Date*, *Depart.Date*, *Entry.Date*, *Max.Birth.Date*, *Min.Birth.Date*, *Entry.Type*, *Depart.Type*, *Sex*, *birthType*, and *AnonInstitutionID*.
* `BirthType`: When separated the analysis by sex category, the birth type to be selected: Captive, Wild, or All
* `xMax` Maximum age in years
* `PlotDir`: Directory to save the plots of the best models

It return a list per sex category including:
* a summary of the data used
* the basta fit of the best model
* the DIC table comparing the different fit of the models
* the remaining life expectancy per age

```{r function-Sur_main}
#' Main survival analysis for taxon profiles
#' 
#' Run the survival analysis per sex and per birth type
#'
#' @param data.core \code{data.frame} including at least the following columns *anonID*, *Birth.Date* (\code{date}), *Depart.Date* (\code{date}), *Entry.Date* (\code{date}), *Max.Birth.Date* (\code{date}), *Min.Birth.Date* (\code{date}), *Entry.Type*, *Depart.Type* *Sex*, *birthType*, and *AnonInstitutionID*
#' @param BirthType \code{character} When separated the analysis by sex category, the birth type to be selected: Captive, Wild, or All Default =  "All"
#' @param outlLev1 \code{numeric} Start threshold used to selected for the data: 100%, 99.9, 99 or 95%
#' @param models \code{vector of characters} names of the basta models to run: "G0", "EX", "LO" and/or "WE". see ?basta for more information. Default = "GO"
#' @param shape \code{character} shape of the basta model: "simple", "Makeham", "bathtub".  see ?basta for more information. Default = "simple"
#' @param mindate \code{character 'YYYY-MM-DD'} Earlier date to include data
#' @param minNsur \code{numeric} Minimum number of individual records needed to run the survival analysis. Default = 50
#' @param minlx  \code{numeric} between 0 and 1. Minimum reached survivorship from the raw Kaplan Meier analysis needed to run the survival analysis. Default = 0.1
#' @param MinBirthKnown  \code{numeric} between 0 and 1. Minimum proportion of individuals with a known birth month in the data. Default = 0.3
#' @param xMax \code{numeric} Maximum age in years Default = 120
#' @param niter  \code{numeric}. number of MCMC iterations to run. see ?basta for more information. Default = 25000
#' @param burnin  \code{numeric} Number of iterations removed so that the model has time to converge. see ?basta for more information. Default = 5001
#' @param thinning  \code{numeric} Number of iteration to run before saving a set of parameters. see ?basta for more information. Default = 20
#' @param nchain  \code{numeric} Number of chains to run. Default = 5001
#' @param ncpus  \code{numeric} Number of computer core to use. Default = 2
#' @param PlotDir \code{character} Directory to save the plots. Default = NULL, no plot is saved
#' @param plotname \code{character} Name used to save the plot. Default = ""
#' 
#' @return The output of a list including per sex categories:
#' * a summary of the data used:
#'- NGlobal: Number of captive born individuals selected from global collections
#'- NBasta: Number of data (individuals) selected for the BaSTA/survival analysis
#'- Ndead:Number of individuals with known age of death used in the BaSTA/survival analysis
#'- maxAge: Maximum age of known age individuals
#'- maxAlive:Maximum number of years spent ex situ
#'- lxMin:Minimum survivorship reached with the raw Kaplan-Meier model
#'-  OutLev: threshold selected for the distribution of  time spent alive: 100%, 99.9, 99 or 95%
#'- a logical indicated if the growth analysis was performed
#'-  If the survival analysis was not performed, an error and its number (Nerr) are returned: The possibility for  this functions are: 1/No raw data and 2/lxMin > minlx 3/NBasta = 0 4/ %known births < MinBirthKnown 5/Data from 1 Institution 6/Nbasta < minNsur, 7/no DIC from Basta.
#'* the basta fit of the best model
#'* the DIC table comparing the different fit of the models
#'* the estimated remaining life expectancy per age
#' If PlotDir is filled, 2 plots are produced: one showing the outliers removed from the data, and one showing the fit of the model on the data.
#'
#' @export
#'
#' @importFrom grDevices pdf dev.off
#'
#' @examples
Sur_main <- function(data.core,  BirthType = "All",
                      models = "GO", shape= "simple", 
                      outlLev1 = NA,
                      mindate = "1980-01-01", minNsur = 50, 
                      minlx = 0.1, MinBirthKnown = 0.3, 
                      xMax = 120,
                      niter = 25000, burnin = 5001, thinning = 20, nchain = 3, 
                      ncpus = 2, PlotDir = NULL, plotname = '') {
  
  mindate = lubridate::as_date(mindate)
  assert_that(is.data.frame(data.core ))
  assert_that(data.core  %has_name% c("anonID", "BirthDate", "DepartDate",
                                      "EntryDate", "MaxBirthDate", "MinBirthDate",
                                      "EntryType", "DepartType", "firstInst", 
                                      "lastInst", "Sex", "birthType"))
  if(!is.na(outlLev1)){
  assert_that(is.numeric(outlLev1))
  assert_that(outlLev1 <= 100)
  }else(outlLev1 = 100)
  assert_that(is.numeric(minNsur))
  assert_that(minNsur > 0)
  assert_that(is.numeric(minlx))
  assert_that(minlx > 0)
  assert_that(minlx < 1)
  assert_that(is.numeric(MinBirthKnown))
  assert_that(MinBirthKnown > 0)
  assert_that(MinBirthKnown <1)
  assert_that(is.numeric(niter))
  assert_that(niter > 0)
  assert_that(is.numeric(burnin))
  assert_that(burnin > 0)
  assert_that(burnin < niter)
  assert_that(is.numeric(thinning))
  assert_that(thinning > 0)
  assert_that(thinning < niter)
  assert_that(is.numeric(nchain))
  assert_that(nchain > 0)
  assert_that(is.numeric(ncpus))
  assert_that(ncpus > 0)
  assert_that(is.character(models))
  assert_that(is.character(plotname))
  assert_that(all(models %in% c("GO", "EX", "LO", "WE")))
  assert_that(is.character(shape))
  assert_that(all(shape %in% c("simple", "bathtub", "Makeham")))
  assert_that(is.numeric(xMax))
  assert_that(xMax > 1)
    assert_that(is.character(BirthType))
  assert_that(length(BirthType) == 1, msg = "You can use only one birth type for each sex analysis")
  assert_that(BirthType %in% c("Captive", "Wild", "All"))
  if(!is.null(PlotDir)){
    checkmate::assert_directory_exists(PlotDir)
  }
  
  # Survival list object:
  surv <- list()
  
  # Run analyses:
    if (BirthType != "All"){
      sexData<- data.core %>%
        filter(stringr::str_detect(birthType, pattern = BirthType))
    }else{sexData<- data.core}

    # Survival Analysis
    out<- Sur_ana(sexData, outlLev1 = outlLev1, models = "GO", shape = "simple",
                   mindate = mindate, minNsur = minNsur, 
                   minlx = minlx, MinBirthKnown = MinBirthKnown, 
                   niter = niter, burnin = burnin, thinning = thinning, nchain = nchain, ncpus = ncpus) 
    out <- append(out, list(relex = NULL))
    if (out$summary$analyzed) {
      
      # Find age at S(x) = 0.001:
      Sx <- out$bastaRes$surv$nocov
      idxMax <- out$bastaRes$x[which(Sx < 0.001)][1]
      if (is.na(idxMax)) {
        xMax <- max(xv)
      } 
      # Remaining life expectancy:
      out$relex <- Sur_relex(theMat = out$bastaRes$params, 
                             model = outBasta$modelSpecs["model"], 
                             shape = shape, ncpus = ncpus, 
                             xMax = xMax, dx = 0.01)
      
      
      #Plots
      if(out$summar$analyzed && !is.null(PlotDir)){
        pdf(file = paste0(PlotDir,"/", plotname, "sur.pdf", sep=""), width = 6, height = 9)
        plot(out$bastaRes)
        plot(out$bastaRes, plot.type = 'demorates')
        plot(out$bastaRes, plot.type = 'gof')
        dev.off()
      }
      
      # Goodness of fit tests
      ##Test if the minimum life expectancy is below 2 years old
      if(min(remex$RemLExp)>2){
        out$summar$error = "Min(Life_exp)>2"
        out$summar$Nerr=8
        out$summar$analyzed <- FALSE 
        out$bastaRes <- list() }
      
      ##Test if the survivorship at mean life expectancy is higher than 0.1
      lx <-outBasta$lifeTable$noCov$Mean$lx
      #age for mle
      dif <-abs(outBasta$lifeTable$noCov$Mean$Ages - outBasta$PS$nocov$PS[1,1])
      
      if(lx[which(dif == min(dif))]< 0.1){
        out$summar$analyzed <- FALSE
        out$summar$error = "lx[MLE]<0.1"
        out$summar$Nerr = 9
        out$bastaRes <- list()   
      }
      
    }
  return(out)
}

```

```{r examples-Sur_main}
data(core)
out <- Sur_main(core, BirthType = "All",
                 models = "GO", shape = "simple",
                 niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
```

```{r tests-Sur_main}
test_that("Sur_main works", {
  data(core)
  out <- Sur_main(core, BirthType = "Captive", 
                  models = "LO", shape = "bathtub",
                  niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
  expect_named(out, c("summary",  "bastaRes", "DICmods", "relex"))

})
```

# Survival Analysis

This functions runs basta models on the data after checking conditions for correct convergence of the models. It includes the following arguments:
* `sexData`including at least the following columns *anonID*, *Birth.Date*, *Depart.Date*, *Entry.Date*, *Max.Birth.Date*, *Min.Birth.Date*, *Entry.Type*, *Depart.Type*, and *AnonInstitutionID*.
* `models` names of the basta models to run: "G0", "EX", "LO" and/or "WE". see ?basta for more information. Default = "GO"
* ` minNsur` Minimum number of individual records needed to run the survival analysis
* `minlx` Minimum reached survivorship from the raw Kaplan Meier analysis needed to run the survival analysis
* `MinBirthKnown` Minimum proportion of individuals with a known birth month in the data

It return a list including:
* a summary of the data used
* the basta fit of the best model
* the DIC table comparing the different fit of the models

```{r function-Sur_ana}
#' Survival Analysis
#' 
#' Run basta models on the data after checking conditions for correct convergence of the models
#'
#' @param sexData \code{data.frame} including at least the following columns *anonID*, *Birth.Date* (\code{date}), *Depart.Date* (\code{date}), *Entry.Date* (\code{date}), *Max.Birth.Date* (\code{date}), *Min.Birth.Date* (\code{date}), *Entry.Type*, *Depart.Type* and *AnonInstitutionID*
#' @param outlLev1 \code{numeric} Start threshold used to selected for the data: 100%, 99.9, 99 or 95%
#' @param models \code{vector of characters} names of the basta models to run: "G0", "EX", "LO" and/or "WE". see ?basta for more information. Default = "GO"
#' @param shape \code{character} shape of the basta model: "simple", "Makeham", "bathtub".  see ?basta for more information. Default = "simple"
#' @param mindate \code{character 'YYYY-MM-DD'} Earlier date to include data
#' @param minNsur \code{numeric} Minimum number of individual records needed to run the survival analysis. Default = 50
#' @param minlx  \code{numeric} between 0 and 1. Minimum reached survivorship from the raw Kaplan Meier analysis needed to run the survival analysis. Default = 0.1
#' @param MinBirthKnown  \code{numeric} between 0 and 1. Minimum proportion of individuals with a known birth month in the data. Default = 0.3
#' @param niter  \code{numeric}. number of MCMC iterations to run. see ?basta for more information. Default = 25000
#' @param burnin  \code{numeric} Number of iterations removed so that the model has time to converge. see ?basta for more information. Default = 5001
#' @param thinning  \code{numeric} Number of iteration to run before saving a set of parameters. see ?basta for more information. Default = 20
#' @param nchain  \code{numeric} Number of chains to run. Default = 5001
#' @param ncpus  \code{numeric} Number of computer core to use. Default = 2
#'
#' @return The output of a list including:
#' * a summary of the data used:
#'- NGlobal: Number of captive born individuals selected from global collections
#'- NBasta: Number of data (individuals) selected for the BaSTA/survival analysis
#'- Ndead:Number of individuals with known age of death used in the BaSTA/survival analysis
#'- maxAge: Maximum age of known age individuals
#'- maxAlive:Maximum number of years spent ex situ
#'- lxMin:Minimum survivorship reached with the raw Kaplan-Meier model
#'-  OutLev: threshold selected for the distribution of  time spent alive: 100%, 99.9, 99 or 95%
#'- a logical indicated if the growth analysis was performed
#'-  If the survival analysis was not performed, an error and its number (Nerr) are returned: The possibility for  this functions are: 1/No raw data and 2/lxMin > minlx 3/NBasta = 0 4/ %known births < MinBirthKnown 5/Data from 1 Institution 6/Nbasta < minNsur, 7/no DIC from Basta.
#'* the basta fit of the best model
#'* the DIC table comparing the different fit of the models
#' 
#' @export
#'
#' @examples
Sur_ana <- function(sexData, outlLev1 = 100, models = "GO", shape = "simple",
                     mindate = "1980-01-01", minNsur = 50, 
                     minlx = 0.1, MinBirthKnown = 0.3, 
                     niter = 25000, burnin = 5001, thinning = 20, nchain = 3, ncpus = 2) {
  
  mindate = lubridate::as_date(mindate)
  assert_that(is.data.frame(sexData))
  assert_that(sexData %has_name% c("anonID", "BirthDate", "DepartDate", "EntryDate", "MaxBirthDate", "MinBirthDate", "EntryType", "DepartType", "firstInst", "lastInst"))
  assert_that(is.numeric(outlLev1))
  assert_that(outlLev1 <= 100)
  assert_that(is.numeric(minNsur))
  assert_that(minNsur > 0)
  assert_that(is.numeric(minlx))
  assert_that(minlx > 0)
  assert_that(minlx <1)
  assert_that(is.numeric(MinBirthKnown))
  assert_that(MinBirthKnown > 0)
  assert_that(MinBirthKnown <1)
  assert_that(is.numeric(niter))
  assert_that(niter > 0)
  assert_that(is.numeric(burnin))
  assert_that(burnin > 0)
  assert_that(burnin < niter)
  assert_that(is.numeric(thinning))
  assert_that(thinning > 0)
  assert_that(thinning < niter)
  assert_that(is.numeric(nchain))
  assert_that(nchain > 0)
  assert_that(is.numeric(ncpus))
  assert_that(ncpus > 0)
  assert_that(is.character(models))
  assert_that(all(models %in% c("GO", "EX", "LO", "WE")))
  assert_that(is.character(shape))
  assert_that(all(shape %in% c("simple", "bathtub", "Makeham")))
  
  
  #Initialize
  summar = list(
    NGlobal =  nrow(sexData), NBasta = 0, Ndead = 0, 
    maxAge = NULL, maxAlive = NULL, 
    lxMin = NULL, outLev = NULL, 
    analyzed = FALSE, Nerr = 0,  error ="")
  bastaRes = NULL
  DICmods = NULL
  #Find the minimum threshold for which lxmin  >.1
  summar$lxMin <- 1
  outLev2 = outlLev1
  while(summar$lxMin > 0.1 & outLev2 >= 95){
    summar$outLev = outLev2
    if (summar$outLev ==100){
      data_sel <-  sexData
    }else{
      data_sel <-  sexData%>%
        filter(!!sym(paste0("above", summar$outLev))==0)
    }
    data_sel <- data_sel%>%
      mutate(
        deparAge = (DepartDate - BirthDate) / 365.25,
        entryAge = (EntryDate - BirthDate) / 365.25
      )
    
    if (!all(data_sel$DepartType == "C")) {
      rawPLE <- Sur_ple(data_sel)
      # Minimum value of lx:
      summar$lxMin <- min(rawPLE$ple)
    }
    if( summar$outLev == 100){outLev2 = 99.9}
    if( summar$outLev == 99.9){outLev2 = 99}
    if( summar$outLev == 99){outLev2 = 95}
    if( summar$outLev == 95){outLev2 = 90}
  }
  
  
  if (summar$lxMin <= minlx) {
    #raw median life expectancy
    # summar$MedLE = median(deparAge[deparType == "D"])
    
    # Extract BaSTA table:
    bastalist <- surv_Bastab(data_sel, earliestDate = mindate,
                               excludeStillbirth = TRUE)
    bastatab <- bastalist%>%
      mutate(
        bdun = Max.Birth.Date-Min.Birth.Date,
        aliveTime = (Depart.Date - Entry.Date) / 365.25)
    
    summar$NGlobal <- nrow(data_sel)
    summar$NBasta <- nrow(bastatab)
    summar$maxAge <- as.numeric(max(bastatab$Depart.Date - bastatab$Birth.Date, na.rm = TRUE))
    summar$maxAlive <- as.numeric(max(bastatab$Depart.Date - bastatab$Entry.Date, na.rm = TRUE))
    
    if(summar$NBasta>0){
      #Check the percentage of individuals with known births
      Perbirthknown =  length(which(bastatab$bdun<32 & bastatab$Entry.Type=="B")) / 
        summar$NBasta
      if(Perbirthknown >= MinBirthKnown){
        #Check that we have more than 1 institution
        Instb =  unique(data_sel$firstInst[data_sel$anonID %in% bastatab$anonID])
        Instl =  unique(data_sel$lastInst[data_sel$anonID %in% bastatab$anonID])
        if(length(unique(c(Instb,Instl)))>1){
          if (summar$NBasta >= minNsur) {
            tempList <- list()
            DICmods <- tibble(models,
                              DIC = 0)
            
            
            for (imod in 1:length(models)) {
              print(models[imod])
              tempList[[models[imod]]] <- BaSTA::basta(
                bastatab, dataType = "census", shape = shape, 
                model = models[imod], parallel = TRUE, 
                ncpus = ncpus, nsim = nchain,
                niter = niter, burnin = burnin, thinning = thinning)
              
              if (!is.na( tempList[[models[imod]]]$DIC[1])) {
                DICmods$DIC[imod] <-  tempList[[models[imod]]]$DIC["DIC"]
              }
            }
            if (all(DICmods$DIC == 0)) {
              print("more chains")
              for (imod in 1:length(models)) {
                print(models[imod])
                tempList[[models[imod]]] <- BaSTA::basta(
                  bastatab, dataType = "census", shape = "bathtub", 
                  model = models[imod], parallel = TRUE, 
                  ncpus = ncpus, nsim = nchain, 
                  niter = niter*4, burnin = burnin*4, thinning = thinning)
                if (!is.na(tempList[[models[imod]]]$DIC[1])) {
                  DICmods$DIC[imod] <-tempList[[models[imod]]]$DIC["DIC"]
                }
                
                
              } 
            }
            
            # BaSTA outputs:
            if (any(DICmods$DIC != 0)) {
              idModSel <- which(DICmods$DIC == min(DICmods$DIC, na.rm = TRUE))
              bastaRes <- tempList[[idModSel]]
            } else {
              summar$error = 'no DIC from Basta'
              summar$Nerr = 7
            }
          } else {
            summar$error = "Nbasta < minNsur"
            summar$Nerr = 6
          }
        }else{
          summar$error = "Data from 1 Institution"
          summar$Nerr = 5
        }
      }else{ 
        summar$error = "%known births < MinBirthKnown"
        summar$Nerr = 4
      }
    }else{
      summar$error = "NBasta = 0"
      summar$Nerr = 3
    }
  } else {
    summar$ error = "lxMin > minlx"
    summar$ Nerr = 2
  }
  
  return(list(summary = summar, bastaRes = bastaRes, DICmods = DICmods))
}
```

```{r examples-Sur_ana}
data(core)
out <- Sur_ana(core,  models = "GO", shape = "simple",
                niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
```

```{r tests-Sur_ana}
test_that("Sur_ana works", {
  data(core)
  out <- Sur_ana(core,  models = "GO", shape = "bathtub",
                  niter = 1000, burnin = 101, thinning = 10, nchain = 3, ncpus = 3)
  expect_named(out, c("summary",  "bastaRes", "DICmods"))
  expect_named(out$DICmods, c("models",  "DIC"))
  expect_equal(nrow(out$DICmods),1)
  expect_named(out$summary, c("NGlobal", "NBasta", "Ndead", "maxAge", "maxAlive", "lxMin",    "outLev", "analyzed", "Nerr", "error"))
  expect_false(out$summary$analyzed )
  expect_true(is.numeric(out$summary$NGlobal ))
  expect_true(is.numeric(out$summary$NBasta ))
  expect_true(is.numeric(out$summary$Ndead ))
  expect_true(is.numeric(out$summary$maxAge ))
  expect_true(is.numeric(out$summary$maxAlive))
  expect_true(is.numeric(out$summary$lxMin ))
  expect_true(is.numeric(out$summary$outLev ))
  expect_true(is.numeric(out$summary$Nerr ))
  expect_true(class(out$bastaRes) == "basta")
})
```



# Create Basta data frame

This functions format the data prior to run basta models. It checks column names and the succession of dates for each individual. It includes the following arguments:

* `coresubset` data frame including at least the following columns *anonID*, *species*, *binSpecies*, *Class*, *Order*, *Family*, *common*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, *DepartType*, and *causeDeath*
* `earliestDate` Earlier date to be included.
* `latestDate` LAtest date to be included.
* `otherCovars` Additional variables to include in the data
* `excludeStillbirth` Whether to exclude still births.

This function return the subset dataset excluding individuals:
* with NA in the columns BirthDate, MinBirthDate, MaxBirthDate, EntryDate, and DepartDate.
* for which the dates from min birth date/ entry date to Depart date do not follow one another.
* Still born if required
* with depart date anterior to earliest date
* with entry date posterior to latest date
* Depart date posterior to latest date are changed to latest date. These individuals are considered as right-censored.



```{r function-surv_Bastab}
#' Create Basta data frame
#' 
#' Check column names and succession of dates to prepare the data for BASTA.
#'
#' @param coresubset  \code{data.frame} including at least the following columns *anonID*, *species*, *binSpecies*, *Class*, *Order*, *Family*, *common*, *BirthDate*, *DepartDate*, *EntryDate*, *MaxBirthDate*, *MinBirthDate*, *EntryType*, *DepartType*, and *causeDeath*
#' @param earliestDate \code{character 'YYYY-MM-DD'} Earlier date to be included. Default = NA
#' @param latestDate \code{character 'YYYY-MM-DD'} LAtest date to be included. Default = NA
#' @param otherCovars \code{vector of character}. Additional variables to include in the data Default = NA
#' @param excludeStillbirth \code{logical} Whether to exclude still births. Default = FALSE
#' 
#' @details
#' This function removes:
#' * individuals with NA in the columns BirthDate, MinBirthDate, MaxBirthDate, EntryDate, and DepartDate.
#' * individuals for which the dates from min birth date/ entry date to Depart date do not follow one another.
#' * Still born individuals if required
#' * individuals with depart date anterior to earliest date
#' * individuals with entry date posterior to latest date
#' * Depart date posterior to latest date are changed to latest date. These individuals are considered as right-censored.
#'
#' @return the subset dataset in the basta format
#' 
#' @export
#'
#' @examples
surv_Bastab <- function (coresubset, earliestDate = NA, latestDate = NA, 
                         otherCovars = NA, excludeStillbirth = FALSE) 
{ 
  inclcols <- c("anonID", "species", "binSpecies", "Class", 
                "Order", "Family", "common", "BirthDate", "MinBirthDate", 
                "MaxBirthDate", "EntryDate", "DepartDate", "EntryType", 
                "DepartType", "causeDeath")
  assert_that(coresubset %has_name% inclcols)
  assert_that(is.logical(excludeStillbirth))
  
  
  if (!is.na(otherCovars)) {
    assert_that(coresubset %has_name% otherCovars)
    inclcols <- c(inclcols, otherCovars)
    ncolnames <- c(ncolnames, otherCovars)
  }
  
  bastadat <- coresubset[, inclcols]
  colnames(bastadat) <- c("anonID", "species", "binSpecies", "Class", 
                          "Order", "Family", "common", "Birth.Date", "Min.Birth.Date", 
                          "Max.Birth.Date", "Entry.Date", "Depart.Date", "Entry.Type", 
                          "Depart.Type", "causeDeath")
  
  
  #Earliest and latest dates
  if (is.na(earliestDate)) {
    earliestDate <- min(bastadat$Min.Birth.Date, na.rm = TRUE)
  }else{
    earliestDate <- lubridate::as_date(earliestDate)
  }
  if (is.na(latestDate)) {
    latestDate <- lubridate::today()
  }else{
    latestDate <- lubridate::as_date(latestDate)
  }
  
  bastadat <- bastadat%>%
    mutate(Depart.Date = lubridate::as_date(Depart.Date),
           Entry.Date = lubridate::as_date(Entry.Date),
           Min.Birth.Date = lubridate::as_date(Min.Birth.Date),
           Max.Birth.Date = lubridate::as_date(Max.Birth.Date),
           Birth.Date = lubridate::as_date(Birth.Date)
    )%>%
    filter(Depart.Date >= earliestDate,
           Entry.Date <= latestDate)%>%
    tidyr::drop_na(c(Birth.Date, Min.Birth.Date, Max.Birth.Date, Entry.Date, Depart.Date))%>%
    filter(Min.Birth.Date <= Birth.Date, 
           Birth.Date <= Max.Birth.Date, 
           Birth.Date <= Entry.Date, 
           Entry.Date <= Depart.Date)%>%
    mutate(Depart.Type = if_else (Depart.Date > latestDate, "C", Depart.Type),
           Depart.Date = if_else (Depart.Date > latestDate, latestDate, Depart.Date),
           Entry.Type  = if_else (Entry.Date < earliestDate, "C", Entry.Type),
           Entry.Date  = if_else (Entry.Date < earliestDate, earliestDate, Entry.Date)
    )
  
  if(excludeStillbirth){
    #Remove Stillborn
    bastadat <- bastadat%>%
      filter(Depart.Date != Birth.Date,
             stringr::str_detect(causeDeath, "Stillborn", negate = T),
             stringr::str_detect(causeDeath, "Fetal death", negate = T))
  }
  
  # class(bastadat) <- "bastazimstab"
  
  return(bastadat)
}
```

```{r examples-surv_Bastab}
data(core)
out<- surv_Bastab(core, earliestDate = '1990-01-01', latestDate = '2020-12-31', 
                  otherCovars = NA, excludeStillbirth = TRUE)


```

```{r tests-surv_Bastab}
test_that("surv_Bastab works", {
  data(core)
  out<- surv_Bastab(core, earliestDate = '1990-01-01', latestDate = '2020-12-31', 
                    otherCovars = NA, excludeStillbirth = TRUE)
  expect_named(out, c("anonID", "species", "binSpecies", "Class", "Order", "Family", "common", "Birth.Date", "Min.Birth.Date", "Max.Birth.Date", "Entry.Date", "Depart.Date", "Entry.Type", "Depart.Type", "causeDeath" ))
  expect_true(min(out$Entry.Date)>= lubridate::as_date('1990-01-01'))
  expect_true(max(out$Depart.Date)<=lubridate::as_date('2020-12-31'))
})
```




# Estimate raw survivorship from life table

This function takes sexData, a data frame including at least the following columns *deparAge*, *entryAge* (\code{date}), and *DepartType*

```{r function-Sur_ple}
#' Estimate raw survivorship from life table
#' 
#' @param sexData \code{data.frame} including at least the following columns *deparAge*, *entryAge* (\code{date}), and *DepartType* 
#'
#' @return A data frame including 3 colummns:
#' * Ages: the different ages
#' * Surv: survivorship
#' * event = 1 if the individual was dead at this age 
#' vs. 0 is the individual was left-censored at this age
#' 
#' @export
#'
#' @examples

# Product limit estimator:
Sur_ple <- function(sexData) {
  assert_that(is.data.frame(sexData))
  assert_that(sexData %has_name% c("deparAge", "DepartType", "entryAge"))
  # Find records with same first and last age:
  sexData <- sexData%>%
    rowwise()%>%
    mutate(idsame = entryAge == deparAge,
           deparAge = ifelse(idsame, deparAge + 1/365.25, deparAge))%>%
    arrange(deparAge)
  
  # Number of ages:
  nage <- nrow(sexData)
  
  # Cx and delta:
  Cx <- rep(0, nage)
  delx <- rep(0, nage)
  
  # Fill up Cx and delta:
  for (ii in 1:nage) {
    agev = sexData$deparAge[ii]
    idNx <-sexData %>%filter(entryAge <= agev,
                             deparAge >= agev)
    Cx[ii] <- nrow(idNx) / nage
    if (sexData$DepartType[ii] == "D") delx[ii] <- 1
  }
  
  # Calculate product limit estimator:
  ple <- cumprod((1 - 1 / (nage * Cx))^delx)
  ple[ple<0]=0
  
  # Add age 0:
  Ages <- sexData$deparAge
  if (Ages[1] > 0) {
    Ages <- c(0, Ages)
    ple <- c(1, ple)
    delx<- c(0, delx)
  }
  
  # Output:
  pleTab <- data.frame(Ages = Ages, ple = ple, event = delx)
  
  return(pleTab)
}

```

```{r examples-Sur_ple}
entryAge = sample(c(1:10), 200, replace = TRUE)
data <- data.frame(
  entryAge = entryAge,
  deparAge =  entryAge + sample(c(0:10), 200, replace = TRUE),
  DepartType = sample(c('C', 'D'), 200, replace = TRUE))


out<-Sur_ple(data)
```

```{r tests-Sur_ple}
test_that("Sur_ple works", {
  entryAge = sample(c(1:10), 200, replace = T)
  data <- data.frame(
    entryAge = entryAge,
    deparAge =  entryAge + sample(c(0:10), 200, replace = T),
    DepartType = sample(c('C', 'D'), 200, replace = T))
  out<-Sur_ple(data)
  expect_named(out, c("Ages", "ple", "event"))
  expect_true(min(out$ple)>=0)
  expect_true(max(out$ple)<=1)
})
```


# Remaining life expectancy

`Sur_relex()` estimate the remaining life expectancy over ages. It includes the following arguments: 

* `theMat` including the posteriors estimates of the model parameter
* `model` names of the basta models to run: "G0", "EX", "LO" and/or "WE".
* `shape` character shape of the basta model: "simple", "Makeham", "bathtub". 
* `xMax` Maximum age in years
* `ncpus`; the number of core to use

```{r function-Sur_relex}
#' Remaining life expectancy
#' 
#' Estimate remaining life expectancy over age from the parameters outputs of a Basta model.
#' 
#'
#' @param theMat \code{array} including the posteriors estimates of the model parameter
#' @param model \code{character} names of the basta models to run: "G0", "EX", "LO" and/or "WE". see ?basta for more information. Default = "GO"
#' @param shape \code{character} shape of the basta model: "simple", "Makeham", "bathtub".  see ?basta for more information. Default = "simple"
#' @param ncpus  \code{numeric} Number of core to use
#' @param xMax \code{numeric} Maximum age in years Default = 120
#' @param dx \code{numeric} precision for age Default = 0.01
#'
#' @return a data frame including age, the mean and 95% credible interval of the remaining life expectancy
#' 
#' @export
#' 
#' @importFrom paramDemo CalcSurv
#' @importFrom snowfall sfInit sfLibrary sfClusterApplyLB  sfStop
#' 
#' @examples
Sur_relex <- function(theMat, model = 'GO', shape = 'bathtub', ncpus = 1,
                      xMax = 120, dx = 0.01) {
  
  assert_that(is.array(theMat))
  assert_that(is.numeric(xMax))
  assert_that(xMax > 1)
  assert_that(is.numeric(dx))
  assert_that(dx > 0)
  assert_that(is.numeric(ncpus))
  assert_that(ncpus > 0)
  assert_that(is.character(model))
  assert_that(all(model %in% c("GO", "EX", "LO", "WE")))
  assert_that(is.character(shape))
  assert_that(all(shape %in% c("simple", "bathtub", "Makeham")))
  
  iseq <- floor(seq(0, nrow(theMat), length = ncpus + 1))
   xv <- seq(0, xMax, by = dx)
 
  # run parallel estimation:
  sfInit(parallel = TRUE, cpus = ncpus)
  # Upload paramDemo:
  # sfLibrary(paramDemo)
  # export variables:
  # sfExport(list = c("iseq", "theMat", "model", "shape", "xMax", "dx"))
  # Run parallel function:
  exparal <- sfClusterApplyLB(1:ncpus, Sur_relex_0, theMat = theMat,
                              model = model, shape = shape,  
                              iseq = iseq, 
                              xMax = xMax, dx = dx, xv = xv )
  
  # Stop application:
  sfStop()
  
  # Gather estimates:
  for (jj in 1:ncpus) {
    if (jj == 1) {
      exMat <- exparal[[jj]]
    } else {
      exMat <- rbind(exMat, exparal[[jj]])
    }
  }
  
  exQuants <- data.frame(Age = xv[which(xv <= xMax)], 
                         RemLExp = apply(exMat, 2, mean), 
                         Lower = apply(exMat, 2, quantile, 0.025),
                         Upper = apply(exMat, 2, quantile, 0.975))
  return(exQuants)
}

```

```{r examples-Sur_relex_parr}
theMat = as.matrix(data.frame( b0 = rnorm(10, -6, 0.01),
                               b1= rnorm(10, 0.1, 0.01)))


out <- Sur_relex(theMat, model = 'GO', shape = 'simple', ncpus = 2,
                 xMax = 50, dx = 0.1)
```

```{r tests-Sur_relex}
theMat = as.matrix(data.frame( b0 = rnorm(10, -6, 0.01),
                               b1= rnorm(10, 0.1, 0.01)))

test_that("Sur_relex works", {
  out <- Sur_relex(theMat, model = 'GO', shape = 'simple', ncpus = 2,
                   xMax = 50, dx = 0.1)
  expect_named(out, c("Age", "RemLExp", "Lower", "Upper"))
  expect_true(is.numeric(out$RemLExp))
})
```

```{r function-Sur_relex_0}
#' Raw remaining life expectancy
#'
#' @return a data frame including age, the mean and 95% credible interval of the remaining life expectancy
#' 
#' @importFrom paramDemo CalcSurv
#' 
#'
#' @examples
#' @noRd
Sur_relex_0 <- function(sim= 1, theMat ,model = 'GO', shape = 'bathtub',  
                        iseq = 1:nrow(theMat), xMax = 120, dx = 0.01 , xv = seq(0, xMax, by = dx)
 ) {
  
  idseq <- (iseq[sim] + 1):iseq[sim + 1]
  
  remex <- t(sapply(idseq, function(ith) {
    theta <- theMat[ith, ]
    Sx <- paramDemo::CalcSurv(theta = theta, x = xv, model = model, shape = shape)
    ex <- Sx * 0
    idn0 <- which(Sx > 0.001)
    ex[idn0] <- rev(cumsum(rev(Sx[idn0] * dx))) / Sx[idn0]
    return(ex)
  }))
  return(remex)
}
```

```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_survival.Rmd", vignette_name = "Survival")
```

